---
title: "Disease Models II"
author: "LRA"
date: "4/17/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rptR)
library(TMB)
library(glmmTMB)
library(bbmle)
library(DHARMa)
library(patchwork)
library(multcomp)
```

## Modeling seagrass wasting disease

Here, I modeled seagrass wasting disease at the meadow scale. Prevalence and severity measurements on individual blades were averaged for each meadow (3702 blades scanned across 32 meadows).

Explanatory variables were Blade Area, Shoot Density, Epiphyte Mass, and Spring Warming.

Spring Warming was derived from MUR SST product - 1 km pixels. Because many of our sites are in small estuaries, the coastal masking of land excludes many sites (11 of 32 meadows). Without in situ HOBO data from 2019 for those sites, I plan to look for other sources of temp data (e.g. NOAA, field station loggers)

Specfically for Spring Warming, I calculated the rate of warming in ºC per week from April 1-June 30 2019.

I also included Region as a fixed effect.

Below is an abbreviated summary of the meadow level modeling with temperature.

```{r data,echo=FALSE}
sp <- read.csv("input data/spring_warming_shoots.csv")
# Center and scale continuous effects
sp$sBladeArea <- scale(sp$BladeAreaMean,center=TRUE,scale=TRUE)
sp$sDensityShoots <- scale(sp$DensityShoots,center=TRUE,scale=TRUE)
sp$sEpiphyteMass <- scale(sp$EpiphyteMass,center=TRUE,scale=TRUE)
sp$sSlope <- scale(sp$slope,center=TRUE,scale=TRUE)
# Since prev and sev at the meadow scale are (0,1), I can use beta regression for both
```

## Prevalence

I used beta regression to model Prevalence as the proportion of infected blades in a meadow as a function of the parameters above.

During initial model selection, the best model had Blade Area, Spring Warming, and Region as predictors. All were significant predictors. Note, "sSlope" is the Spring Warming effect (scaled slope of temp increase in Spring).

Post-hoc contrasts showed that the Regions grouped into "Super Regions" 

AK and BC were in one group (not significantly different form each other) and WA, OR, SD were in another group (not significantly different from each other and different from the AK/BC group). These Super Regions followed geographic patterns, so I re-grouped the meadows as Northern (AK and BC) and Southern (WA, OR, SD). Note, BB sites were excluded for lack of temperature data. 

```{r prev1, echo=FALSE}
fit_prev <- glmmTMB(PrevalenceMean~sBladeArea+sSlope+Region,
                     data=sp,
                     family=beta_family(link = "logit"))
summary(fit_prev)

glht_glmmTMB <- function (model, ..., component="cond") {
  glht(model, ...,
       coef. = function(x) fixef(x)[[component]],
       vcov. = function(x) vcov(x)[[component]],
       df = NULL)
}
modelparm.glmmTMB <- function (model, coef. = function(x) fixef(x)[[component]],
                               vcov. = function(x) vcov(x)[[component]],
                               df = NULL, component="cond", ...) {
  multcomp:::modelparm.default(model, coef. = coef., vcov. = vcov.,
                               df = df, ...)
}
g1 <- glht(fit_prev,linfct=mcp(Region="Tukey"))
summary(g1)
```


With the new Super Region grouping, I re-ran the model with the same other effects (Blade Area and Spring Warming)

```{r prev2,echo=FALSE}
sp$SuperRegion[sp$Region=="AK"|sp$Region=="BC"] <- "Northern"
sp$SuperRegion[sp$Region=="WA"|sp$Region=="OR"|sp$Region=="SD"] <- "Southern"

fit_prev2_new <- glmmTMB(PrevalenceMean~sBladeArea+sSlope+SuperRegion,
                     data=sp,
                     family=beta_family(link = "logit"))
summary(fit_prev2_new)
```

In the Super Region model, Spring Warming and Super Region were highly significant. This suggests there is a temperature effect!

The plots below show a model visualization - predicted values of wasting disease prevalence at the meadow scale for each Super Region. The relationships are quite distinct. Lines show the model predictions, points show the empirical data.

```{r prev_vis,echo=FALSE}
jvalues <- with(sp, seq(from = min(slope), to = max(slope), length.out = 500))
sjvalues <- scale(jvalues,center=TRUE,scale=TRUE)
b.data <- data.frame(sSlope=sjvalues,sBladeArea=-0.3808)
superregions <- c("Northern","Southern")
for(i in seq_along(superregions)){
  b.data$SuperRegion <- rep.int(superregions[i],nrow(b.data))
  # b.data$Pred <- predict(fit_prev,newdata=b.data,type='response')
  if(i==1){dat2 <- b.data}else{dat2 <- rbind(dat2,b.data)}
}
pred2 <- predict(fit_prev2_new,newdata=dat2,se=TRUE,type='response')
try2 <- as.data.frame(pred2)
try2 <- within(try2, {
  PredictedInfection <- (fit)
  LL <- (fit - (1.96 * se.fit))
  UL <- (fit + (1.96 * se.fit))
})
try2 <- cbind(try2,dat2)
try2$Slope <- (try2$sSlope*sd(jvalues))+mean(jvalues)

northern <- subset(try2,SuperRegion=="Northern")
southern <- subset(try2,SuperRegion=="Southern")

Np <- ggplot(northern,aes(x=Slope))+
  geom_line(aes(y=PredictedInfection))+
  geom_line(aes(y=LL),linetype="dashed")+
  geom_line(aes(y=UL),linetype="dashed")+
  geom_point(data=sp[sp$SuperRegion=="Northern",],aes(x=slope,y=PrevalenceMean))+
  ylim(c(-0.003,1.01))+
  theme_bw()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Northern Meadows",
       subtitle = "n=11 meadows, British Columbia and Alaska")

Sp <- ggplot(southern,aes(x=Slope))+
  geom_line(aes(y=PredictedInfection))+
  geom_line(aes(y=LL),linetype="dashed")+
  geom_line(aes(y=UL),linetype="dashed")+
  geom_point(data=sp[sp$SuperRegion=="Southern",],aes(x=slope,y=PrevalenceMean))+
  ylim(c(-0.003,1.01))+
  theme_bw()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Southern Meadows",
       subtitle = "n=10 meadows, San Diego to Washington")

Np+Sp
```

A few caveats. Clearly, the data set is limited - we only have 21 meadows. And, especially for the Northern meadows, they are bunched in one part of the distribution. Adding the remaining 11 sites, if we can get workable temperature data, might improve the model - or it might totally wreck it, if BB sites do not follow this pattern! 

## Severity

Finally, for severity, I used the same approach (beta regression, same predictors)

I am still working on model parameterization here - I'm not totally happy with the model yet. But, the significant predictors and therefore model inference have been pretty consistent as I'm tweaking the model: Blade Area is a (weakly) significant predictor of severity at the meadow level. Spring Warming is not. 

There's no pattern of Super Regions in the Region contrasts with Severity.

```{r sev,echo=FALSE}
fit_sev1 <- glmmTMB(SeverityMean~sBladeArea+sSlope+Region,
                    data=sp,
                    family=beta_family(link = "logit"))
summary(fit_sev1)
g2 <- glht(fit_sev1,linfct=mcp(Region="Tukey"))
summary(g2)
```

Overall, it looks like the rate of warming in spring may be an important predictor for wasting disease prevalence, and that there is a strong geographic effect. Geographic pattern could related to divergent populations, phenotypic plasticity in response to different warming or other environmental factors... 