---
title: "RelativeTempDisease"
author: "LRA"
date: "9/24/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
region_order <- c("AK","BC","WA","OR","BB","TB","SD")
Month_order <- c("January","February","March","April","May","June","July","August","September","October","November","December")
meadow_order <- c("AK_A","AK_B","AK_C","AK_D","AK_E","AK_F",
                  "BC_A","BC_B","BC_C","BC_D","BC_E",
                  "WA_A","WA_B","WA_C","WA_D","WA_E",
                  "OR_A","OR_B","OR_C","OR_D","OR_E",
                  "BB_A","BB_B","TB_C","BB_D","TB_E","TB_F",
                  "SD_A","SD_B","SD_C","SD_D","SD_E")
my_mean <- function(x) mean(x[x != 0],na.rm = TRUE)
```

### Quick summary of relative temperature metrics

Here I am trying to model the effects of temperature on disease across the latitudinal scale. There's no relationship between any absolute temperature metric and disease levels. 

Proposed metric is number of hot days in Summer (July and August) that are hot within the region

```{r data}
daily <- read.csv("input data/daily_supp.csv")
disease <- read.csv("input data/disease_2019.csv")

disease_site <- disease %>%
  group_by(Region,SiteCode)%>%
  summarise(SeverityMean=my_mean(Severity),
            PrevalenceMean=mean(Prevalence),
            BladeAreaMean=mean(BladeArea),
            LesionAreaMean=mean(LesionArea))
# disease_site$Region <- as.character(disease_site$Region)
# disease_site$Region[disease_site$Region=="BB"&disease_site$SiteCode=="C"] <- "TB"
# disease_site$Region[disease_site$Region=="BB"&disease_site$SiteCode=="E"] <- "TB"
# disease_site$Region[disease_site$Region=="BB"&disease_site$SiteCode=="F"] <- "TB"
# disease_site$Region <- as.factor(disease_site$Region)
disease_site$Region <- ordered(disease_site$Region,levels=c("AK","BC","WA","OR","BB","SD"))

disease_site$Meadow <- paste(disease_site$Region,disease_site$SiteCode,sep="_")

daily$date <- as.POSIXct(daily$date)
daily$week <- floor_date(daily$date,unit="week",week_start = getOption("lubridate.week.start", 7))
daily$month <- floor_date(daily$week,unit="month")
daily$month.w <- as.character.Date(daily$month,format="%B")
# Define seasons by patterns in temp - constant low temps through March, rising temps through June, pretty stable during July/Aug
daily$season[daily$month.w=="January" | daily$month.w== "February"|daily$month.w== "March"|daily$month.w=="December"] <- "Winter"
daily$season[daily$month.w=="June" | daily$month.w== "April"|daily$month.w== "May"] <- "Spring"
daily$season[daily$month.w== "July" | daily$month.w== "August"] <- "Summer"
daily$region <- as.character(daily$region)
daily$region[daily$region=="BB"&daily$site=="C"] <- "TB"
daily$region[daily$region=="BB"&daily$site=="E"] <- "TB"
daily$region[daily$region=="BB"&daily$site=="F"] <- "TB"
daily$region <- as.factor(daily$region)
daily$region <- ordered(daily$region,levels=c("AK","BC","WA","OR","BB","TB","SD"))
daily$meadow <- paste(daily$region,daily$site,sep="_")
daily$meadow <- ordered(daily$meadow,levels=meadow_order)
regional <- daily %>%
  group_by(region,season,month.w,month,week,date)%>%
  summarise(RegionTemp=mean(analysed_sst,na.rm = TRUE))
ggplot()+
  geom_line(data=daily,aes(x=date,y=analysed_sst,color=site))+
  geom_line(data=regional,aes(x=date,y=RegionTemp),linetype="dashed")+
  facet_wrap(~region)+
  theme_bw()

ggplot()+
  geom_line(data=daily[daily$season=="Summer",],aes(x=date,y=analysed_sst,color=site))+
  geom_line(data=regional[regional$season=="Summer",],aes(x=date,y=RegionTemp),linetype="dashed")+
  facet_wrap(~region,scales="free")+
  theme_bw()

regional_stats <- regional %>%
  group_by(region,season)%>%
  summarise(min=min(RegionTemp),
            fifth=quantile(RegionTemp,0.05),
            tenth=quantile(RegionTemp,0.10),
            Q1=quantile(RegionTemp,0.25),
            med=median(RegionTemp),
            mean=mean(RegionTemp),
            Q3=quantile(RegionTemp,0.75),
            eighty=quantile(RegionTemp,0.8),
            ninety=quantile(RegionTemp,0.9),
            ninetyfive=quantile(RegionTemp,0.95),
            max=max(RegionTemp),
            range=max-min)
print(regional_stats)
```

## HotDays

Comparing disease levels based on the number of days above the 90th percentile within a region

```{r threshold, echo=FALSE,warning=FALSE}
# Only calculating the warm threshold here (HotDays) because ColdDays isn't as informative
meadows <- levels(daily$meadow)
regions <- levels(daily$region)
sites <- levels(daily$site)
winter <- subset(regional_stats,season=="Winter")
summer <- subset(regional_stats,season=="Summer")
alldat <- data.frame()
for(i in seq_along(meadows)){
  summer_days <- subset(daily,season=="Summer"&meadow==meadows[i])
  region <- summer_days$region[1]
  threshold <- summer$ninety[summer$region==region]
  hot_days <- length(summer_days$date[summer_days$analysed_sst>threshold])
  dat <- data.frame(Meadow=meadows[i],Threshold=threshold,HotDays=hot_days)
  if(i==1){alldat <- dat}else{alldat <- rbind(alldat,dat)}
}
ninety <- na.omit(alldat)
ninety <- left_join(ninety,disease_site,by="Meadow")
ggplot(ninety,aes(x=HotDays,y=PrevalenceMean,color=Region))+geom_point()+
  theme_bw()
ggplot(ninety,aes(x=HotDays,y=PrevalenceMean,color=Region))+geom_point()+
  geom_smooth(method="lm",se=FALSE)+
  theme_bw()
```

This is a relative temperature metric that puts all the sites on the same temperature scale. But it is relative within regeions, and would be a lot better based on long-term climatology versus one summer. 

Also, BB and TB only have 3 sites each, which is pretty limiting. 