---
title: "MURTempExp"
author: "LRA"
date: "4/17/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Data exploration with SST from MUR data product 

# Daily SST 

```{r daily,echo=FALSE}
region_order <- c("AK","BC","WA","OR","BB","SD")
Month_order <- c("January","February","March","April","May","June","July","August","September","October","November","December")
daily <- read.csv("input data/daily_MUR_temps.csv")
monthly <- read.csv("input data/monthly_MUR_temps.csv")
sat_loc <- read.csv("input data/sat_locations.csv")
disease <- read.csv("input data/disease_2019.csv")

daily$date <- as.POSIXct(daily$date)
daily$month <- as.POSIXct(daily$month)
daily$month.w <- as.character.Date(daily$month,format="%B")
daily$season[daily$month.w=="January" | daily$month.w== "February"|daily$month.w== "March"] <- "Winter"
daily$season[daily$month.w=="April" | daily$month.w== "May"|daily$month.w== "June"] <- "Spring"
daily$season[daily$month.w=="July" | daily$month.w== "August"] <- "Summer"
daily$season <- ordered(daily$season,levels=c("Winter","Spring","Summer"))
daily$region <- ordered(daily$region,levels=c("AK","BC","WA","OR","BB","SD"))

ggplot(daily,aes(x=date,y=analysed_sst,color=site))+geom_line()+
  facet_wrap(~region)+
  theme_bw()+
  xlab("Date")+
  ylab("Daily SST ºC)")+
  labs(title="Daily SST from MUR product, Jan-Aug 2019")
```

Key caveat - all BB sites are offshore, hence the cooler and less seasonal temperatures are not accurate for seagrass sites. 
Two OR and three SD sites are also offshore. We do not have any HOBO data for this period (Jan-Aug 2019) but we may be able to find some buoy or other data to supplement for at least some BB sites.

If we plot at weekly intervals, we can see seasonal patterns more clearly.
```{r weekly,echo=FALSE}
daily$week <- floor_date(daily$date,unit="week",week_start = getOption("lubridate.week.start", 7))

weekly <- daily %>%
  group_by(month,week,region,site,latitude,longitude) %>%
  summarise(weeklySST=mean(analysed_sst))

ggplot(weekly,aes(x=week,y=weeklySST,color=site))+geom_point()+
  facet_wrap(~region)+
  theme_bw()+
  xlab("Date")+
  ylab("Weekly SST ºC)")+
  labs(title="Weekly SST from MUR product, Jan-Aug 2019")

```

Focus on spring (April, May June) as a period of rapid temperature increase at most sites, with temps plateauing during later summer (except maybe in SD).

Also, eliminate meadows that are offshore (11) leaving n=21 meadows in the remaining data set.

## Spring temps
```{r spring, echo=FALSE,warning=FALSE}
spring <- subset(weekly,month>"2019-03-31" & month<"2019-07-01")
spring <- left_join(spring,sat_loc,by=c("region"="Region","site"="SiteCode"))
spring <- subset(spring,SatLoc==TRUE)
spring$fweek <- as.factor(spring$week)
spring$nweek <- as.integer(spring$fweek)
spring$region <- ordered(spring$region,levels=region_order)

spring$meadow <- as.factor(paste(spring$region,spring$site,sep="_"))
meadows <- levels(spring$meadow)

for(i in seq_along(meadows)){
  mod <- lm(weeklySST~nweek,data=spring[spring$meadow==meadows[i],])
  slope <- mod$coefficients[2]
  r2 <- as.character(signif(summary(mod)$adj.r.squared, 2))
  p <- as.character(signif(summary(mod)$coef[2,4], 2))
  dat <- data.frame(meadow=meadows[i],slope=slope,r2=r2,p=p)
  if(i==1){alldat <- dat}else{alldat <- rbind(alldat,dat)}
}

disease_site <- disease %>%
  group_by(Region,SiteCode)%>%
  summarise(SeverityMean=mean(Severity),
            PrevalenceMean=mean(Prevalence),
            BladeAreaMean=mean(BladeArea),
            LesionAreaMean=mean(LesionArea))
disease_site$Region <- ordered(disease_site$Region,levels=c("AK","BC","WA","OR","BB","SD"))

disease_site$Meadow <- paste(disease_site$Region,disease_site$SiteCode,sep="_")
alldat <- left_join(alldat,disease_site,by=c("meadow"="Meadow"))

ggplot(spring,aes(x=week,y=weeklySST,color=site))+geom_point()+
  facet_wrap(~region)+
  theme_bw()+
  xlab("Date")+
  ylab("Weekly SST ºC)")+
  labs(title="Weekly SST from MUR product, April-June 2019")

ggplot(alldat,aes(x=slope,y=PrevalenceMean,color=Region))+geom_point()+
  theme_bw()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Disease prevalence as a function of spring warming")

ggplot(alldat,aes(x=slope,y=SeverityMean,color=Region))+geom_point()+
  theme_bw()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease severity (% leaf area)")+
  labs(title="Disease severity as a function of spring warming")


```

There's clear differentiation in spring warming rates between regions, and between sites within regions.
WA.C is Beach Haven - warms up really fast and is the highest disease in the project.

Fairly nice relationship between spring warming rate and prevalence and severity. It would be really good to get the 11 missing values on these plots.

## Seasonal change
We can also look at temps by month to see if there are strong patterns in any month. 

```{r monthly,echo=FALSE}
monthly$month <- as.POSIXct(monthly$month)
monthly$month.w <- as.character.Date(monthly$month,format="%B")
monthly$region <- ordered(monthly$region,levels=c("AK","BC","WA","OR","BB","SD"))
monthly <- left_join(monthly,disease_site,by=c("region"="Region","site"="SiteCode"))
monthly$month.w <- ordered(monthly$month.w,levels=Month_order)
ggplot(monthly,aes(x=SSTmean,y=SeverityMean,color=region))+geom_point()+
  facet_wrap(~month.w)+
  theme_bw()+
  xlab("Mean SST (ºC)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Disease prevalence as a function of monthly SST")

ggplot(monthly,aes(x=SSTmean,y=PrevalenceMean,color=region))+geom_point()+
  facet_wrap(~month.w)+
  theme_bw()+
  xlab("Mean SST (ºC)")+
  ylab("Wasting disease severity (% leaf area)")+
  labs(title="Disease severity as a function of monthly SST")
```

No relationships between disease metrics and monthly mean temperatures. 

However, as an alternative to the spring warming rate, we can look at the change from Winter (Jan-March) to Spring (April-June) temps.

```{r seasonal,echo=FALSE,warning=FALSE}
seasonal <- daily %>%
  group_by(region,site,latitude,longitude,season)%>%
  summarise(SSTseason=mean(analysed_sst))
seasonal_wide <- seasonal[,c(1,2,5,6)]
seasonal_wide <- spread(seasonal_wide,key=season,value=SSTseason)
seasonal_wide$dWS <- seasonal_wide$Spring-seasonal_wide$Winter
seasonal_wide$dSS <- seasonal_wide$Summer-seasonal_wide$Spring
seasonal_wide$dWSum <- seasonal_wide$Summer-seasonal_wide$Winter
seasonal_wide1 <- left_join(seasonal_wide,disease_site,by=c("region"="Region","site"="SiteCode"))
seasonal_wide4 <- left_join(seasonal_wide1,sat_loc,by=c("region"="Region","site"="SiteCode"))
seasonal_wide4 <- subset(seasonal_wide4,SatLoc=="TRUE")
seasonal_wide4$region <- ordered(seasonal_wide4$region,levels=region_order)

ggplot(seasonal_wide4,aes(x=dWS,y=PrevalenceMean,color=region))+geom_point()+
  theme_bw()+
  xlab("Increase in mean temp from Winter to Spring (ºC)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Disease prevalence as a function of spring warming")
## this one looks good
ggplot(seasonal_wide4,aes(x=dWS,y=SeverityMean,color=region))+geom_point()+
  theme_bw()+
   xlab("Increase in mean temp from Winter to Spring (ºC)")+
  ylab("Wasting disease severity (% leaf area)")+
  labs(title="Disease severity as a function of spring warming")
```

A couple observations: BC is somewhat out of the general pattern between the other sites. Relationship is more clear for prevalence, whereas severity, without Beach Haven (WA.C), the relationship is minimal.