---
title: "SpringWarmingModel"
author: "LRA"
date: "9/29/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(betareg)
library(performance)
```

### Spring warming model

Here I use beta regression to model disease prevalence as a function of spring warming rates, i.e. weekly increase in temperature from April-June

Temperature data are mostly from MUR SST (n=21 meadows) but some are from in situ sources (n=5)

BB_E is excluded - if BB_E is included, there is no significant effect of the model.

Here, I'm showing a final model that separates out two superregions - AK and BC together and other sites are grouped together.

Note the main difference between these regions is due to colder winters in AK and BC - they warm up faster to get to similar temps to WA/OR in summer. 

```{r data}
region_order <- c("AK","BC","WA","OR","BB","SD")
alldat <- read.csv("input data/new_spring_warming.csv")
alldat$Region <- ordered(alldat$Region,levels=region_order)
```


```{r model, echo=FALSE}
alldat$SuperRegion[alldat$Region=="AK"|alldat$Region=="BC"] <- "N"
alldat$SuperRegion[alldat$Region=="WA"|alldat$Region=="OR"|alldat$Region=="BB"|alldat$Region=="SD"] <- "S"

fit4 <- betareg(PrevalenceMean~slope+SuperRegion,
                data=alldat[alldat$meadow!="BB_E",],
                link="logit")
summary(fit4)
plot(fit4)
jvalues <- with(alldat, seq(from = min(slope), to = max(slope), length.out = 500))
sjvalues <- scale(jvalues,center=TRUE,scale=TRUE)
b.data <- data.frame(slope=jvalues)
superregions <- c("N","S")

out <- data.frame()
for(i in seq_along(superregions)){
  b.data$SuperRegion <- rep.int(superregions[i],nrow(b.data))
  if(i==1){out <- b.data}else{out <- rbind(out,b.data)}
}
pred <- cbind(
  response=predict(fit4,newdata=out,type='response'),
  variance=predict(fit4,newdata=out,type='variance'),
  predict(fit4,newdata=out,type='quantile',at=c(0.025,0.975)))
predict(fit4)
try <- as.data.frame(pred)
try <- cbind(try,out)
northern <- subset(try,SuperRegion=="N")
southern <- subset(try,SuperRegion=="S")

ggplot(northern,aes(x=slope))+
  geom_line(aes(y=response),color="grey50")+
#  geom_line(aes(y=q_0.025),linetype="dashed",color="grey50")+
 # geom_line(aes(y=q_0.975),linetype="dashed",color="grey50")+
  geom_point(data=alldat[alldat$meadow!="BB_E",],aes(x=slope,y=PrevalenceMean,color=Region))+
  geom_line(data=southern,aes(x=slope,y=response),color="grey15")+
  #geom_line(data=southern,aes(x=slope,y=q_0.025),linetype="dashed",color="grey15")+
  #geom_line(data=southern,aes(x=slope,y=q_0.975),linetype="dashed",color="grey15")+
  geom_point(data=alldat[alldat$meadow=="BB_E",],aes(x=slope,y=PrevalenceMean),shape=8,color="darkred")+
  theme_bw()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Diseaes prevalence as a function of spring warming",
       subtitle = "Temperatures from MUR SST and in situ, n=26 meadows")

ggplot(northern,aes(x=slope))+
  geom_line(aes(y=response),color="grey50")+
  geom_line(aes(y=q_0.025),linetype="dashed",color="grey50")+
  geom_line(aes(y=q_0.975),linetype="dashed",color="grey50")+
  geom_point(data=alldat[alldat$meadow!="BB_E",],aes(x=slope,y=PrevalenceMean,color=Region))+
  geom_line(data=southern,aes(x=slope,y=response),color="grey15")+
  geom_line(data=southern,aes(x=slope,y=q_0.025),linetype="dashed",color="grey15")+
  geom_line(data=southern,aes(x=slope,y=q_0.975),linetype="dashed",color="grey15")+
  geom_point(data=alldat[alldat$meadow=="BB_E",],aes(x=slope,y=PrevalenceMean),shape=8,color="darkred")+
  theme_bw()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease prevalence (% infected)")+
  labs(title="Diseaes prevalence as a function of spring warming",
       subtitle = "Temperatures from MUR SST and in situ, n=26 meadows")

```
```{r severity}
fit5 <- betareg(SeverityMean~slope,
                data=alldat[alldat$meadow!="BB_E",],
                link="logit")
summary(fit5)
plot(fit5)
r2(fit5)

jvalues <- with(alldat, seq(from = min(slope), to = max(slope), length.out = 100))
b.data <- data.frame(slope=jvalues)

pred <- cbind(
  response=predict(fit5,newdata=b.data,type='response'),
  variance=predict(fit5,newdata=b.data,type='variance'),
  predict(fit5,newdata=b.data,type='quantile',at=c(0.025,0.975)))
try <- as.data.frame(pred)
try <- cbind(try,b.data)
ggplot(try,aes(x=slope))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=alldat,aes(x=slope,y=SeverityMean,color=Region))+
  scale_color_viridis_d()+
  xlab("Slope of spring warming (ºC per week)")+
  ylab("Wasting disease severity (% blade area)")+
  labs(title="Diseaes severity as a function of spring warming",
       subtitle = "Temperatures from MUR SST and in situ, n=26 meadows")+
  theme_bw()
```