---
title: "PredictingTomalesBayTemps"
author: "LRA"
date: "4/28/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)

```

### Using BOON buoy temperatures to predict site temperatures at Tomales Bay seagrass meadows

Analysis here uses three sources of temperature data:

(1) logger temperatures from Jay's students from 2017-May 2019 at the Tomales Bay meadows

(2) buoy temperatures from the Tomales Bay buoy, maintained by BML, from 2017-2019

(3) NSF HOBO logger temperatures from July-Aug 2019

I used the buoy temps and logger temps from the overlapping period 2017-May 2019 to calculate regression models to predict site temperatures.

Then, I predicted site temperatures during the data gap (May 16-July 4 2019) using the buoy data.

One question is how comparable the other loggers and the NSF HOBO loggers are - if they are subtidal vs intertidal, could change the relationship with the buoys.

```{r data,echo=FALSE}
# Bring in logger temperature data
tmls <- read.csv("input data/supplemental temps/TomalesBayTempsJune2017toApril2019.csv")
tmls$Date <- as.POSIXct(tmls$Date, format="%Y-%m-%d %H:%M:%S")

# Calculate hourly logger temperatures
tmls_hourly <- tmls %>%
  group_by(Site,Year,DateTime=floor_date(Date,unit="hour")) %>%
  summarise(LoggerHourly=mean(Temp,na.rm=TRUE))

# Bring in buoy data
tbb_17 <- read.csv("input data/supplemental temps/tbb_seawater_temperature_2017_hourly.csv",col.names=c("","time","","temp"))
tbb_17 <- tbb_17[,c("time","temp")]
tbb_17$time <- as.POSIXct(tbb_17$time, format="%Y-%m-%d %H:%M:%S")
# Drop low values for when the buoy was not recording
tbb_17$temp[tbb_17$temp<5] <- NA

tbb_18 <- read.csv("input data/supplemental temps/tbb_seawater_temperature_2018_hourly.csv",col.names = c("","time","","temp"))
tbb_18 <- tbb_18[,c("time","temp")]
tbb_18$time <- as.POSIXct(tbb_18$time,format="%Y-%m-%d %H:%M:%S")
tbb_18$temp[tbb_18$temp<5] <- NA

tbb_19 <- read.csv("input data/supplemental temps/tbb_seawater_temperature_2019_hourly.csv",col.names = c("","time","","temp"))
tbb_19 <- tbb_19[,c("time","temp")]
tbb_19$time <- as.POSIXct(tbb_19$time,format="%Y-%m-%d %H:%M:%S")
tbb_19$temp[tbb_19$temp<5] <- NA

tbb <- rbind(tbb_17,tbb_18,tbb_19)
tbb <- na.omit(tbb)

tbb_hourly <- tbb %>%
  group_by(DateTime=floor_date(time,unit="hour"))%>%
  summarise(BuoyHourly=mean(temp,na.rm=TRUE))

combo <- left_join(tmls_hourly,tbb_hourly,by="DateTime")
combo$Year <- as.factor(combo$Year)
combo <- na.omit(combo)
# Model site temp (logger temp) as a function of buoy temp
BL <- subset(combo,Site=="Blakes")
MP <- subset(combo,Site=="Millerton")
NC <- subset(combo,Site=="Nicks")
BL_mod <- lm(LoggerHourly~BuoyHourly,data=BL)
MP_mod <- lm(LoggerHourly~BuoyHourly,data=MP)
NC_mod <- lm(LoggerHourly~BuoyHourly,data=NC)
```

### Model summaries for site temp vs buoy temp

All three models are significant, R2 ~0.6-0.7

```{r model, echo=FALSE,warning=FALSE}
# Summaries of the site level models predicting site temperature from buoy temperature
summary(BL_mod)
summary(MP_mod)
summary(NC_mod)
```

```{r model_plot,echo=FALSE}
sBL <- coefficients(BL_mod)[2]
iBL <- coefficients(BL_mod)[1]
BL$PredictedHourly <- iBL+BL$BuoyHourly*sBL

sMP <- coefficients(MP_mod)[2]
iMP <- coefficients(MP_mod)[1]
MP$PredictedHourly <- iMP+MP$BuoyHourly*sMP

sNC <- coefficients(NC_mod)[2]
iNC <- coefficients(NC_mod)[1]
NC$PredictedHourly <- iNC+NC$BuoyHourly*sNC

combo2 <- rbind(BL,MP,NC)

pred_2019 <- subset(combo2,Year=="2019")

pred_2019_daily <- pred_2019 %>%
  group_by(Site,Year,DateTime=floor_date(DateTime,unit="day"))%>%
  summarise(LoggerDaily=mean(LoggerHourly,na.rm=TRUE),
            BuoyDaily=mean(BuoyHourly,na.rm=TRUE),
            PredictedDaily=mean(PredictedHourly,na.rm=TRUE))
ggplot(pred_2019_daily,aes(x=LoggerDaily,y=PredictedDaily,color=Year))+geom_point()+
  facet_wrap(~Site)+
  geom_abline()+
  theme_bw()+
  xlab("Logger Daily Temperature (ºC)")+
  ylab("Predicted Daily Temperature (ºC)")

# ggplot(pred_2019_daily,aes(x=DateTime))+geom_point(aes(y=LoggerDaily),color="darkblue")+
#   geom_point(aes(y=PredictedDaily),color="lightblue")+
#   facet_wrap(~Site)+
#   geom_abline()+
#   theme_bw()
```

The predicted temperatures match well with the logger temperatures from Jan-May 2019. However, there's more spread (divergence from the 1:1 line) at higher temperatures.

```{r data2, echo=FALSE,warning=FALSE}
# Read in NSF logger data
hobo <- read.csv("input data/supplemental temps/BB.alltemps.2019.csv")
hobo$DateObs <- as.POSIXct(hobo$DateObs,format=("%m/%d/%y %H:%M"))

hobo_tmls <- subset(hobo,SiteCode=="C"|SiteCode=="E"|SiteCode=="F")
hobo_tmls <- hobo_tmls[,c("SiteCode","SiteName","DateObs","TempC")]
hobo_tmls_daily <- hobo_tmls %>%
  group_by(SiteCode,SiteName,DateTime=floor_date(DateObs,unit="day"))%>%
  summarise(TempDaily=mean(TempC,na.rm = TRUE))
hobo_2019 <- subset(hobo_tmls_daily,DateTime>"2019-07-04")
hobo_2019$Site <- recode(hobo_2019$SiteName,"Blake's Landing"="Blakes","Millerton Point"="Millerton","Nick's Cove"="Nicks")

# NSF HOBO temps start on 7/4 and 7/5 - use 7/5 for all three for simplicity

tbb_daily <- tbb %>%
  group_by(DateTime=floor_date(time,unit="day"))%>%
  summarise(BuoyDaily=mean(temp,na.rm=TRUE))
tbb_daily_19 <- subset(tbb_daily,DateTime>"2018-12-31")
tbb_hourly_gap <- subset(tbb_hourly,DateTime>"2019-05-15" & DateTime<"2019-07-05")

buoy_2019 <- bind_rows("Blakes"=tbb_hourly_gap,"Millerton"=tbb_hourly_gap,"Nicks"=tbb_hourly_gap,.id="Site")

buoy_2019$Slope <- NA
buoy_2019$Intercept <- NA
buoy_2019$Slope[buoy_2019$Site=="Blakes"] <- sBL
buoy_2019$Slope[buoy_2019$Site=="Millerton"] <- sMP
buoy_2019$Slope[buoy_2019$Site=="Nicks"] <- sNC

buoy_2019$Intercept[buoy_2019$Site=="Blakes"] <- iBL
buoy_2019$Intercept[buoy_2019$Site=="Millerton"] <- iMP
buoy_2019$Intercept[buoy_2019$Site=="Nicks"] <- iNC

buoy_2019$PredictedHourly <- buoy_2019$Intercept+buoy_2019$BuoyHourly*buoy_2019$Slope
buoy_2019 <- buoy_2019 %>%
  group_by(Site,DateTime=floor_date(DateTime,unit="day"))%>%
  summarise(PredictedDaily=mean(PredictedHourly,na.rm=TRUE))

logger_2019 <- pred_2019_daily[,c("Site","DateTime","LoggerDaily")]
buoy_2019 <- plyr::rename(buoy_2019,replace=c("PredictedDaily"="TempDaily"))
logger_2019 <- plyr::rename(logger_2019,replace=c("LoggerDaily"="TempDaily"))
full_2019 <- bind_rows(logger_2019,buoy_2019,hobo_2019)

full_2019$DateTime <- as.Date.POSIXct(full_2019$DateTime)
tbb_daily_19$DateTime <- as.Date.POSIXct(tbb_daily_19$DateTime)
```

### Comparing the 2019 temperature record

```{r combo,echo=FALSE,warning=FALSE}
ggplot(full_2019,aes(x=DateTime,y=TempDaily,color=Site))+geom_line()+
  geom_line(data=tbb_daily_19[tbb_daily_19$DateTime<"2019-09-01",],aes(x=DateTime,y=BuoyDaily,color="TB buoy"))+
  geom_vline(xintercept=as.numeric(as.Date(c("2019-05-15","2019-07-05"))),linetype=4)+
  theme_bw()+
  ylab("Daily temperature (ºC)")+
  xlab("Date")+
  labs(title="Combined Tomales Bay temperatures 2019")
```

Temperatures from Jan 1 - May 15 are from logger data (from Jay's other loggers)

Temperatures between the dashed lines (May 16 - July 4) are predicted from the buoy temperatures.

Temperatures from July 5 - Aug 31 are from the NSF HOBO loggers.

Predictions are made at the hourly level and averaged for each day.

Some issues: 

(1) During the prediction period, there's more separation between Millerton and the other sites. 

(2) Millerton is the hottest site consistently through the prediction period but not during the other periods. 

(3) Steep drop in temps in mid-May (i.e. moving from the loggers to predicted temps).

But, the July temperatures don't have a big jump when moving back from predictions to loggers, and the spread between the buoy and site temps during the prediction period seems consistent with the July-Aug period.

A couple other notes:

I tried modeling site temps from buoy temps during only the "gap" (May-June 2017-2019). However, the linear relationship between hourly site temps and hourly buoy temps from just the May-June period is really weak (R2 <0.1). I think this suggests that there's a lot of variability in spring that means the buoy temp is a worse predictor of site-level temps than during other periods (e.g. winter). That's not very surprising, but is unfortunate if we want to use spring temperatures as a predictor of disease.

An alternative to predicting site temps from buoy temps is to accept the gap in the temp data (May 16-July 4) at the Tomales sites. If we have permission to uses these other loggers (from Katie DuBois?), we have _some_ spring and summer temps, so we can use that more limited dataset for the 2019 disease modeling. 