---
title: "Blade_models_hurdle_glmm"
author: "LRA"
date: "1/26/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(patchwork)
library(performance)
library(sjPlot)
```

## Blade level models of prevalence and severity

Blade level models are fit as hurdles - binomial/logistic regression for prevalence, beta regression for severity of diseased blades only

Two sets of models, one with cumulative positive temperature anomaly as a predictor, one without.

Because only 27 meadows out of 32 have SST data for the anomaly calculation, the number of replicates is smaller for those models.

However, the significant effects are mainly the same across the models.

Predictors are: Blade Area, Shoot Density, Epiphyte Mass per Blade Area, Cumulative positive anomaly, Tidal Height, and all interactions between Tidal Height and fixed effects.

Random structure for blades is Transect within Meadow within Region. However, for the CPTA models (models that include the temperature anomaly as a predictor), Region is not sufficiently variable to be used as a random effect (possibly due to unbalanced availability of SST across regions). Consistently, across all models, Region is not a significant fixed effect either and the model can be simplified to remove Region. 

Final CPTA models therefore use Transect within Meadow for the random structure at the blade level. 

Here I am not showing all the detail of model selection and validation (some things are commented out) so that the summary outputs are from the final models only.

```{r data}
bladeWD <- read.csv("input data/bladeWD.csv")
bladeWD <- bladeWD[,c("Prevalence","Severity","BladeArea","DensityShoots","TidalHeight","Lat","SampleId","Region","SiteCode","Transect")]

epi <- read.csv("Refined Models/input/shoots_epiphytes_summ1_SD.csv")
epi <- select(epi,c("Region","SiteCode","TidalHeight","Transect","EpiphytePerAreaMean"))
bladeWD <- left_join(bladeWD,epi,by=c("Region","SiteCode","TidalHeight","Transect"))
cpta <- read.csv("Refined Models/input/june19_9y_SST_anomaly_disease.csv")
cpta <- select(cpta,c("Region","Site","Meadow","CDiffMeanHeat"))
bladeWD <- left_join(bladeWD,cpta,by=c("Region","SiteCode"="Site"))
bladeWD$sBladeArea <- scale(bladeWD$BladeArea,center=TRUE,scale=TRUE)
bladeWD$sDensityShoots <- scale(bladeWD$DensityShoots,center=TRUE,scale=TRUE)
bladeWD$sEpiphytePerAreaMean <- scale(bladeWD$EpiphytePerAreaMean,center=TRUE,scale=TRUE)
bladeWD$sCDiffMeanHeat <- scale(bladeWD$CDiffMeanHeat,center=TRUE,scale=TRUE)
bladeWD$MeadowId <- paste(bladeWD$Region,bladeWD$SiteCode,sep=".")
bladeWD$TransectId <- paste(bladeWD$MeadowId,bladeWD$Transect,sep="_")
bladeWD$Region <- as.factor(bladeWD$Region)
dat <- bladeWD[-which(is.na((bladeWD$CDiffMeanHeat))),]
#rescale continuous predictors based on the restricted dataset
dat$sBladeArea <- scale(dat$BladeArea,center=TRUE,scale=TRUE)
dat$sDensityShoots <- scale(dat$DensityShoots,center=TRUE,scale=TRUE)
dat$sEpiphytePerAreaMean <- scale(dat$EpiphytePerAreaMean,center=TRUE,scale=TRUE)
dat$sCDiffMeanHeat <- scale(dat$CDiffMeanHeat,center=TRUE,scale=TRUE)

```

### blade level prevalence with CPTA (n=3177) ####

```{r blade_prev_CPTA_selection, echo=FALSE}
# fit_prev1 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                        (1|Region)+(1|MeadowId)+(1|TransectId),
#                     data=dat,
#                     family=binomial(link="logit"))
# summary(fit_prev1)
# E1.sim <- simulateResiduals(fit_prev1)
# plot(E1.sim)
# plot(E1.sim$scaledResiduals~dat$sBladeArea)
# plot(E1.sim$scaledResiduals~dat$sDensityShoots)
# plot(E1.sim$scaledResiduals~as.factor(dat$TidalHeight))
# plot(E1.sim$scaledResiduals~dat$Region)
# drop1(fit_prev1)
# model_performance(fit_prev1)
## so region again isn't the source of variation
#icc(fit_prev1,by_group = TRUE)
# refit with Region as a fixed effect
# fit_prev2 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                        Region+(1|MeadowId)+(1|TransectId),
#                      data=dat,
#                      family=binomial(link="logit"))
# drop1(fit_prev2)
# dropping region improves AIC marginally
fit_prev3 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
#summary(fit_prev3)
#model_performance(fit_prev3)
#icc(fit_prev3,by_group = TRUE)
# grouping is about 2/3 by meadow, 1/3 by transect


### Model selection by AIC
 # remove blade area : tidal height interaction
 fit_prev3.1 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sDensityShoots:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
 
 # remove shoot density : tidal height interaction
fit_prev3.2 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeArea:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
  # remove epiphyte load : tidal height interaction
fit_prev3.3 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                       sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
 # remove cumulative anomaly : tidal height interaction
 fit_prev3.4 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
 # no interactions
 fit_prev3.5 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
 
 # no tidal height (or interactions)
 fit_prev3.6 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
 df.AIC <- AIC(fit_prev3,fit_prev3.1,fit_prev3.2,fit_prev3.3,fit_prev3.4,
     fit_prev3.5,fit_prev3.6)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
 # By AIC and weight, fit_prev3.3 is the best fit
```

By AIC and weight, fit_prev3.3 (no epiphyte load: tidal height interaction) is best.

```{r blade_prev_CPTA_output, echo=FALSE,fig.width=4,fig.height=4}
# Assess the model residuals 
E3.sim <- simulateResiduals(fit_prev3.3)
plot(E3.sim)
plot(E3.sim$scaledResiduals~dat$sBladeArea)
plot(E3.sim$scaledResiduals~dat$sDensityShoots)
plot(E3.sim$scaledResiduals~dat$sEpiphytePerAreaMean)
plot(E3.sim$scaledResiduals~as.factor(dat$TidalHeight))
plot(E3.sim$scaledResiduals~dat$sCDiffMeanHeat)

# Look at the model output, performance metrics
summary(fit_prev3.3)
performance(fit_prev3.3)
## Visualize the best-fitting model output
p_cpta_names <- c("Cumulative SST anomaly * Tidal height (upper)",
                  "Shoot density * Tidal height (upper)", "Leaf area * Tidal height (upper)",
                  "Tidal height (upper)", "Cumulative SST anomaly","Epiphyte load","Shoot density","Blade area")
p_cpta_plot <- plot_model(fit_prev3.3,
           axis.labels = p_cpta_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.4,
            #transform = NULL
           #axis.lim = c(0.1,10),
           
           group.terms = c(1,1,2,1,2,2,1,2)
           )
p_cpta_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("black","grey50"))+
  ylab("Scaled estimates of \ndisease prevalence odds ratio")+
  labs(tag="A")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))
```
### Blade level prevalence, no CPTA n = 3702 #### 

This is the prevalence model without any temperature terms, with the full 3702 blade dataset.

For this model, Region can be included in the random part

```{r blade_prev_all_selection}
fit_prev5 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
#summary(fit_prev5)


### Model selection by AIC
 # remove blade area : tidal height interaction
 fit_prev5.1 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                       sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
 
 # remove shoot density : tidal height interaction
fit_prev5.2 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                       sBladeArea:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
  # remove epiphyte load : tidal height interaction
fit_prev5.3 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
 # no interactions
 fit_prev5.4 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
 # no tidal height (or interactions)
 fit_prev5.5 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
 df.AIC <- AIC(fit_prev5,fit_prev5.1,fit_prev5.2,fit_prev5.3,fit_prev5.4,
     fit_prev5.5)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
By AIC and weight, best-fitting model is the full model (fit_prev5).

```{r blade_prev_all_output, echo=FALSE,fig.width=4,fig.height=4}
# Assess the model residuals 
E5.sim <- simulateResiduals(fit_prev5)
plot(E5.sim)
plot(E5.sim$scaledResiduals~bladeWD$sBladeArea)
plot(E5.sim$scaledResiduals~bladeWD$sDensityShoots)
plot(E5.sim$scaledResiduals~bladeWD$sEpiphytePerAreaMean)
plot(E5.sim$scaledResiduals~as.factor(bladeWD$TidalHeight))

# Look at the model output, performance metrics
summary(fit_prev5)
performance(fit_prev5)
## Visualize the best-fitting model output
p_all_names <- c("Epiphyte load * Tidal height (upper)", "Shoot density * Tidal height (upper)", 
                 "Blade area * Tidal height (upper)","Tidal height (upper)",
                 "Epiphyte load","Shoot density","Blade area")
p_all_plot <- plot_model(fit_prev5,
           axis.labels = p_all_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.4,
            #transform = NULL
           #axis.lim = c(0.1,10),
           
           group.terms = c(1,1,2,2,1,1,2)
           )
p_all_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("black","grey50"))+
  ylab("Scaled estimates of \ndisease prevalence odds ratio")+
  labs(tag="C")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))
```

## Blade level severity, with CPTA (n=1573) ####

Severity at the blade level is the second part of the hurdle, so it is only modeled on the diseased blades.

Same approach to having two models, one with fewer replicates because of the availability of SST for the anomaly calculation.

These models are fit with beta regression, as severity is a non-count proportion.

```{r blade_sev_cpta_selection}
sev <- subset(bladeWD,Prevalence==1)
sev$sBladeArea <- scale(sev$BladeArea,center=TRUE,scale=TRUE)
sev$sDensityShoots <- scale(sev$DensityShoots,center=TRUE,scale=TRUE)
sev$sEpiphytePerAreaMean <- scale(sev$EpiphytePerAreaMean,center=TRUE,scale=TRUE)
sev$sCDiffMeanHeat <- scale(sev$CDiffMeanHeat,center=TRUE,scale=TRUE)

write.csv(sev,"Refined Models/input/severity_blades.csv",row.names=FALSE)
sevdat <- sev[-which(is.na((sev$CDiffMeanHeat))),]
sevdat$sBladeArea <- scale(sevdat$BladeArea,center=TRUE,scale=TRUE)
sevdat$sDensityShoots <- scale(sevdat$DensityShoots,center=TRUE,scale=TRUE)
sevdat$sEpiphytePerAreaMean <- scale(sevdat$EpiphytePerAreaMean,center=TRUE,scale=TRUE)
sevdat$sCDiffMeanHeat <- scale(sevdat$CDiffMeanHeat,center=TRUE,scale=TRUE)

# use beta regression in glmmTMB

# fit_sev1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                       (1|Region)+(1|MeadowId)+(1|TransectId),
#                                data=sevdat,
#                                family=beta_family(link = "logit"))
# summary(fit_sev1)
# # Region doesn't work as a random effect, try as fixed
# fit_sev2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                       Region+(1|MeadowId)+(1|TransectId),
#                     data=sevdat,
#                     family=beta_family(link = "logit"))
# summary(fit_sev2)
# drop1(fit_sev2)
# drop Region as a fixed effect
fit_sev3 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
#summary(fit_sev3)
#performance(fit_sev3)
#icc(fit_sev3,by_group = TRUE)

 ### Model selection by AIC
 # remove blade area : tidal height interaction
 fit_sev3.1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sDensityShoots:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
 
 # remove shoot density : tidal height interaction
fit_sev3.2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeArea:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
  # remove epiphyte load : tidal height interaction
fit_sev3.3 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
 # remove cumulative anomaly : tidal height interaction
 fit_sev3.4 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
 # no interactions
 fit_sev3.5 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
 # no tidal height (or interactions)
 fit_sev3.6 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
 df.AIC <- AIC(fit_sev3,fit_sev3.1,fit_sev3.2,fit_sev3.3,fit_sev3.4,
     fit_sev3.5,fit_sev3.6)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
By AIC and weight, best model is fit_sev3.4 (no cumulative anomlay:tidal height interaction)

```{r blade_sev_CPTA_output, echo=FALSE,fig.width=4,fig.height=4}
# Assess the model residuals 
Es.sim <- simulateResiduals(fit_sev3.4)
plot(Es.sim)
plot(Es.sim$scaledResiduals~sevdat$sBladeArea)
plot(Es.sim$scaledResiduals~sevdat$sDensityShoots)
plot(Es.sim$scaledResiduals~sevdat$sEpiphytePerAreaMean)
plot(Es.sim$scaledResiduals~as.factor(sevdat$TidalHeight))
plot(Es.sim$scaledResiduals~sevdat$sCDiffMeanHeat)
testDispersion(Es.sim)
# Some issues with dispersion here, but we don't need to worry too much for a beta distribution
# Adding a dispersion formula (variable dispersion) doesn't improve the diagnostics
# The residuals plotted against model covariate don't show patterns and have pretty good uniformity, so keep the 3.4 model
# fit_sev_disp <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
#                         sEpiphytePerAreaMean:TidalHeight+
#                       (1|MeadowId)+(1|TransectId),
#                       dispformula = ~sBladeArea,
#                     data=sevdat,
#                     family=beta_family(link = "logit"))
# Edisp.sim <- simulateResiduals(fit_sev_disp)
# plot(Edisp.sim)
# summary(fit_sev_disp)

# Look at the model output, performance metrics
summary(fit_sev3.4)
performance(fit_sev3.4)

## Visualize the best-fitting model output
s_cpta_names <- c("Epiphyte load * Tidal height (upper)",
                  "Shoot density * Tidal height (upper)", "Blade area * Tidal height (upper)",
                  "Tidal height (upper)", "Cumulative SST anomaly","Epiphyte load","Shoot density","Blade area")
s_cpta_plot <- plot_model(fit_sev3.4,
           axis.labels = s_cpta_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.4,
            #transform = NULL
           #axis.lim = c(0.1,10),
           
           group.terms = c(1,1,1,2,2,2,1,2)
           )
s_cpta_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("black","grey50"))+
  ylab("Scaled estimates of \ndisease severity odds ratio")+
  labs(tag="B")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))
```

## blade level severity, no CPTA, n=1853 ####

Follow up model includes all diseased blades, no temperature anomaly term.

Also fit with beta regression.

```{r blade_sev_all}
# fit_sev1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
#                       (1|Region)+(1|MeadowId)+(1|TransectId),
#                     data=sev,
#                     family=beta_family(link = "logit"))
# summary(fit_sev1)
# performance(fit_sev1)
# drop1(fit_sev1)
# Region does work as a random effect
fit_sev2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
#summary(fit_sev2)
#performance(fit_sev2)
#icc(fit_sev2,by_group = TRUE)


### Model selection by AIC
 # remove blade area : tidal height interaction
 fit_sev2.1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
 
 # remove shoot density : tidal height interaction
fit_sev2.2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
  # remove epiphyte load : tidal height interaction
fit_sev2.3 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
 # no interactions
 fit_sev2.4 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
 # no tidal height (or interactions)
 fit_sev2.5 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
 df.AIC <- AIC(fit_sev2,fit_sev2.1,fit_sev2.2,fit_sev2.3,fit_sev2.4,
     fit_sev2.5)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
By AIC and weight, best model is 2.1, no blade area:tidal height interaction

```{r blade_sev_all_output, echo=FALSE,fig.width=4,fig.height=4}
# Assess the model residuals 
E2.sim <- simulateResiduals(fit_sev2.1)
plot(E2.sim)
plot(E2.sim$scaledResiduals~sev$sBladeArea)
plot(E2.sim$scaledResiduals~sev$sDensityShoots)
plot(E2.sim$scaledResiduals~sev$sEpiphytePerAreaMean)
plot(E2.sim$scaledResiduals~as.factor(sev$TidalHeight))
# Again as with blade severity cpta model, there's an issue with underdispersion, but adding a variable dispersion formula 
# doesn't improve residual diagnostics, and the residuals look fine plotted against model covariates. So, accept model.
# Look at the model output, performance metrics
summary(fit_sev2.1)
performance(fit_sev2.1)
## Visualize the best-fitting model output
s_all_names <- c("Epiphyte load * Tidal height (upper)", "Shoot density * Tidal height (upper)", 
                 "Tidal height (upper)",
                 "Epiphyte load","Shoot density","Blade area")
s_all_plot <- plot_model(fit_sev2.1,
           axis.labels = s_all_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.4,
            #transform = NULL
           #axis.lim = c(0.1,10),
           
           group.terms = c(1,1,1,2,1,1)
           )
s_all_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("black","grey50"))+
  ylab("Scaled estimates of \ndisease severity odds ratio")+
  labs(tag="D")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))
```
