---
title: "Blade_models_hurdle_glmm"
author: "LRA"
date: "1/26/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(patchwork)
library(performance)
```

## Blade level models of prevalence and severity

Blade level models are fit as hurdles - binomial/logistic regression for prevalence, beta regression for severity of diseased blades only

Two sets of models, one with cumulative positive temperature anomaly as a predictor, one without.

Because only 27 meadows out of 32 have SST data for the anomaly calculation, the number of replicates is smaller for those models.

However, the significant effects are mainly the same across the models.

Predictors are: Blade Area, Shoot Density, Epiphyte Mass per Blade Area, Cumulative positive anomaly, Tidal Height, and all interactions between Tidal Height and fixed effects.

Random structure for blades is Transect within Meadow within Region. However, for the CPTA models (models that include the temperature anomaly as a predictor), Region is not sufficiently variable to be used as a random effect (possibly due to unbalanced availability of SST across regions). Consistently, across all models, Region is not a significant fixed effect either and the model can be simplified to remove Region. 

Final models therefore use Transect within Meadow for the random structure at the blade level. 

Here I am not showing all the detail of model selection and validation (some things are commented out) so that the summary outputs are from the final models only.

```{r data}
bladeWD <- read.csv("input data/bladeWD.csv")
bladeWD <- bladeWD[,c("Prevalence","Severity","BladeArea","DensityShoots","TidalHeight","Lat","SampleId","Region","SiteCode","Transect")]

epi <- read.csv("Refined Models/input/shoots_epiphytes_summ1_SD.csv")
epi <- select(epi,c("Region","SiteCode","TidalHeight","Transect","EpiphytePerAreaMean"))
bladeWD <- left_join(bladeWD,epi,by=c("Region","SiteCode","TidalHeight","Transect"))
cpta <- read.csv("Refined Models/input/june19_9y_SST_anomaly_disease.csv")
cpta <- select(cpta,c("Region","Site","Meadow","CDiffMeanHeat"))
bladeWD <- left_join(bladeWD,cpta,by=c("Region","SiteCode"="Site"))
bladeWD$sBladeArea <- scale(bladeWD$BladeArea,center=TRUE,scale=TRUE)
bladeWD$sDensityShoots <- scale(bladeWD$DensityShoots,center=TRUE,scale=TRUE)
bladeWD$sEpiphytePerAreaMean <- scale(bladeWD$EpiphytePerAreaMean,center=TRUE,scale=TRUE)
bladeWD$sCDiffMeanHeat <- scale(bladeWD$CDiffMeanHeat,center=TRUE,scale=TRUE)
bladeWD$MeadowId <- paste(bladeWD$Region,bladeWD$SiteCode,sep=".")
bladeWD$TransectId <- paste(bladeWD$MeadowId,bladeWD$Transect,sep="_")
bladeWD$Region <- as.factor(bladeWD$Region)
dat <- bladeWD[-which(is.na((bladeWD$CDiffMeanHeat))),]

```

### blade level prevalence with CPTA (n=3177) ####

```{r blade_prev_CPTA, echo=FALSE}
# fit_prev1 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                        (1|Region)+(1|MeadowId)+(1|TransectId),
#                     data=dat,
#                     family=binomial(link="logit"))
# summary(fit_prev1)
# E1.sim <- simulateResiduals(fit_prev1)
# plot(E1.sim)
# plot(E1.sim$scaledResiduals~dat$sBladeArea)
# plot(E1.sim$scaledResiduals~dat$sDensityShoots)
# plot(E1.sim$scaledResiduals~as.factor(dat$TidalHeight))
# plot(E1.sim$scaledResiduals~dat$Region)
# drop1(fit_prev1)
# model_performance(fit_prev1)
## so region again isn't the source of variation
#icc(fit_prev1,by_group = TRUE)
# refit with Region as a fixed effect
# fit_prev2 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                        Region+(1|MeadowId)+(1|TransectId),
#                      data=dat,
#                      family=binomial(link="logit"))
# drop1(fit_prev2)
# dropping region improves AIC marginally
fit_prev3 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId)+(1|TransectId),
                     data=dat,
                     family=binomial(link="logit"))
summary(fit_prev3)
model_performance(fit_prev3)
icc(fit_prev3,by_group = TRUE)
# grouping is about 2/3 by meadow, 1/3 by transect
# if we take the transect part out
# fit_prev4 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                        (1|MeadowId),
#                      data=dat,
#                      family=binomial(link="logit"))
# summary(fit_prev4)
# model_performance(fit_prev4)
## fit_prev4 is worse by AIC than fit_prev3 so go with fit_prev3
# adjusted alpha = 0.0042
## sCDiffMeanHeat and sBladeArea are significant
# sCDiffMeanHeat is 2.3x greater than Blade Area, so warmer conditons may be more important

```

### Blade level prevalence, no CPTA n = 3702 #### 

This is the prevalence model without any temperature terms, with the full 3702 blade dataset.

For this model, Region can be included in the random part

```{r blade_prev_all}
fit_prev5 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       (1|Region)+(1|MeadowId)+(1|TransectId),
                     data=bladeWD,
                     family=binomial(link="logit"))
summary(fit_prev5)
#drop1(fit_prev5)
# no improvement to the model by dropping non-significant terms
# if k = 11, 0.05/11=0.0045, then blade area is the only significant term
# coefficient of sBladeArea is similar in CPTA/nonCPTA models, so that's encouraging
model_performance(fit_prev5)
icc(fit_prev5,by_group=TRUE)
# fit_prev6 <- glmmTMB(Prevalence~sBladeArea+sDensityShoots+TidalHeight+
#                        sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
#                        (1|Region)+(1|MeadowId)+(1|Transect),
#                      data=bladeWD,
#                      family=binomial(link="logit"))
# anova(fit_prev5,fit_prev6)
```

## Blade level severity, with CPTA (n=1573) ####

Severity at the blade level is the second part of the hurdle, so it is only modeled on the diseased blades.

Same approach to having two models, one with fewer replicates because of the availability of SST for the anomaly calculation.

These models are fit with beta regression, as severity is a non-count proportion.

```{r blade_sev_cpta}
sev <- subset(bladeWD,Prevalence==1)
write.csv(sev,"Refined Models/input/severity_blades.csv",row.names=FALSE)
sevdat <- sev[-which(is.na((sev$CDiffMeanHeat))),]

# use beta regression in glmmTMB

# fit_sev1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                       (1|Region)+(1|MeadowId)+(1|TransectId),
#                                data=sevdat,
#                                family=beta_family(link = "logit"))
# summary(fit_sev1)
# # Region doesn't work as a random effect, try as fixed
# fit_sev2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                     sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                       Region+(1|MeadowId)+(1|TransectId),
#                     data=sevdat,
#                     family=beta_family(link = "logit"))
# summary(fit_sev2)
# drop1(fit_sev2)
# drop Region as a fixed effect
fit_sev3 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|MeadowId)+(1|TransectId),
                    data=sevdat,
                    family=beta_family(link = "logit"))
summary(fit_sev3)
performance(fit_sev3)
icc(fit_sev3,by_group = TRUE)
# Es.sim <- simulateResiduals(fit_sev3)
# plot(Es.sim)
# plot(Es.sim$scaledResiduals~sevdat$sBladeArea)
# plot(Es.sim$scaledResiduals~sevdat$sDensityShoots)
# plot(Es.sim$scaledResiduals~as.factor(sevdat$TidalHeight))
# plot(Es.sim$scaledResiduals~sevdat$Region)
# drop1(fit_sev3)
# do get a warning with the ICC that estimate of random effect may be unreliable - model is too complex?
# try fitting a model with no interactions
# fit_sev4 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
#                     (1|MeadowId)+(1|TransectId),
#                     data=sevdat,
#                     family=beta_family(link = "logit"))
# summary(fit_sev4)
 # performance(fit_sev4)
 # icc(fit_sev4,by_group = TRUE)
 # anova(fit_sev3,fit_sev4)
# model3 can possibly be simplified by dropping 1 or more interaction terms using likelihood test
# AIC doesn't change much though
# so, keep model 3? significance of terms is the same in both
# for model sev3, shoot density is sig and the interaction between shoot density and tidal height
# adjusted alpha is 0.0042

# fit_sev5 <- glmmTMB(Severity~sDensityShoots+sBladeArea+sCDiffMeanHeat+
#                       (1|MeadowId)+(1|TransectId),
#                     data=sevdat,
#                     family=beta_family(link = "logit"))
# summary(fit_sev5)
# something to consider is that the simpler model suggests sBladeArea matters more than sDensityShoots hmm
```

## blade level severity, no CPTA, n=1853 ####

Follow up model includes all diseased blades, no temperature anomaly term.

Also fit with beta regression.

```{r blade_sev_all}
# fit_sev1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
#                       (1|Region)+(1|MeadowId)+(1|TransectId),
#                     data=sev,
#                     family=beta_family(link = "logit"))
# summary(fit_sev1)
# performance(fit_sev1)
# drop1(fit_sev1)
# Region does work as a random effect
fit_sev2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|Region)+(1|MeadowId)+(1|TransectId),
                    data=sev,
                    family=beta_family(link = "logit"))
summary(fit_sev2)
performance(fit_sev2)
icc(fit_sev2,by_group = TRUE)
#anova(fit_sev2,fit_sev1)
# no sig difference so drop the interaction
#drop1(fit_sev2)
# no other terms will improve model by leaving
# Es.sim <- simulateResiduals(fit_sev2)
# plot(Es.sim)
# plot(Es.sim$scaledResiduals~sev$sBladeArea)
# plot(Es.sim$scaledResiduals~sev$sDensityShoots)
# plot(Es.sim$scaledResiduals~as.factor(sev$TidalHeight))
# plot(Es.sim$scaledResiduals~sev$Region)
```

