---
title: "Meadow_scale_model"
author: "LRA"
date: "1/26/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(betareg)
library(performance)
library(patchwork)
library(cowplot)
library(sjPlot)
region_order <- c("AK","BC","WA","OR","BB","SD")

```

### Meadow-scale drivers of wasting disease

Refined model of disease prevalence and severity at the meadow scale

Major predictor is cumulative temperature anomaly as a fixed effect
Other predictors to possibly include are density and blade area at the meadow scale (as fixed effects)
However, replication is n=27, so the number of parameters that can be estimated is small (for a conservative n/k=10, can only fit
one fixed effect due to estimating intercept and dispersion parameter. Up to 2 fixed effects keeps n/k at 6.75, still reasonable)

```{r data}
# read in and arrange data
dat <- read.csv("Refined Models/input/june19_9y_SST_anomaly_disease.csv")
dat$Region <- ordered(dat$Region,levels=region_order)
dat$CPTempAnomaly <- dat$CDiffMeanHeat
ba <- read.csv("Refined Models/input/Measured_Disease_summ4.csv")
ba$Meadow <- paste(ba$Region,ba$SiteCode,sep="_")
ba <- select(ba,c("Meadow","BladeAreaMean"))
dat <- left_join(dat,ba,by="Meadow")
den <- read.csv("Refined Models/input/density_cover_summ4.csv")
den$Meadow <- paste(den$Region,den$SiteCode,sep="_")
den <- select(den,c("Meadow","DensityShootsMean"))
dat <- left_join(dat,den,by="Meadow")
```

### Prevalence model

Compare four models, one with cumulative anomaly and blade area as fixed effects, one with cumulative anomaly and shoot density,
third with only cumulative anomaly and last with all three fixed effects

```{r prev_model_build, echo=TRUE}
fit_prev1 <- betareg(PrevalenceMean~CPTempAnomaly + BladeAreaMean,
                     data=dat,
                     weights = Count,
                     link = "logit")

fit_prev2 <- betareg(PrevalenceMean~CPTempAnomaly + DensityShootsMean,
                     data=dat,
                     weights = Count,
                     link = "logit")
fit_prev3 <- betareg(PrevalenceMean~CPTempAnomaly,
                     data=dat,
                     weights = Count,
                     link = "logit")
fit_prev4 <- betareg(PrevalenceMean~ BladeAreaMean,
                     data=dat,
                     weights = Count,
                     link = "logit")
fit_prev5 <- betareg(PrevalenceMean~ DensityShootsMean,
                     data=dat,
                     weights = Count,
                     link = "logit")

df.AIC <- AIC(fit_prev1,fit_prev2,fit_prev3,fit_prev4,fit_prev5)
df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
First model (cumulative anomaly and blade area) is substantially better by AIC.

```{r prev_model_eval,echo=FALSE,fig.width=4,fig.height=4}
summary(fit_prev1)
E.prev <- resid(fit_prev1,type='pearson')
F.prev <- fitted(fit_prev1)
plot(E.prev~F.prev)
plot(E.prev~dat$BladeAreaMean)
plot(E.prev~dat$CPTempAnomaly)

## Visualize the best-fitting model effect sizes
prev_names <- c("Leaf area", "Cumulative SST anomaly")
meadow_plot <- plot_model(fit_prev1,
                          type="std",
           axis.labels = prev_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.1,
           axis.lim = c(0.2,5),
           group.terms = c(1,1)
           )
meadow_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.5,2.4))+
  scale_color_manual(values=c("black"))+
  ylab("Scaled estimates of \ndisease prevalence odds ratio")+
  labs(tag="A")+
  #xlab("Scaled parameters")+
  theme(panel.grid = element_blank(),
        axis.text = element_text(size=10))

```

No major issues with the residuals when plotted against covariates and fitted values. Model is acceptable.

Both cumulative anomaly and blade area are significant. 

Odds ratio plot shows the standardized effect sizes. For a 1 SD increase in Cumulative SST anomaly, the chances of a meadow being completely diseased (prevalence = 100%) effectively doubles. Effect of leaf area is much weaker (only increases prevalence by a factor of 1.14x for each increase in SD).

Note, SD of Cumulative SST anomaly is 7.4ºC, so for every ~7 degrees of anomaly in the month of June, get a doubling of prevalence

Visualize model by simulating fit across new data holding one variable constant at a time.

```{r prev_model_viz,echo=FALSE,fig.width=5,fig.height=7}
# First simulate change in CPTA
jvaluesPa <- with(dat, seq(from = min(CPTempAnomaly), to = max(CPTempAnomaly), length.out = 100))
# create new data and hold blade area at the median value of the dataset (29.1 cm2)
b.dataPa <- data.frame(CPTempAnomaly=jvaluesPa,BladeAreaMean=median(dat$BladeAreaMean))
predPa <- cbind(
  response=predict(fit_prev1,newdata=b.dataPa,type='response'),
  variance=predict(fit_prev1,newdata=b.dataPa,type='variance'),
  predict(fit_prev1,newdata=b.dataPa,type='quantile',at=c(0.025,0.975)))
preva <- as.data.frame(predPa)
preva <- cbind(preva,b.dataPa)
a <- ggplot(preva,aes(x=CPTempAnomaly))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=CPTempAnomaly,y=PrevalenceMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab("Cumulative positive temperature anomaly (ºC)")+
  ylab("Wasting disease prevalence\n (% individuals infected)")+
  theme_bw()+
  theme(panel.grid = element_blank())

# Next, repeat simulation but hold CPTA constant at median value (11.7 ºC)
jvaluesPb <- with(dat, seq(from = min(BladeAreaMean), to = max(BladeAreaMean), length.out = 100))
# create new data and hold blade area at the median value of the dataset (29.1 cm2)
b.dataPb <- data.frame(BladeAreaMean=jvaluesPb,CPTempAnomaly=median(dat$CPTempAnomaly))
predPb <- cbind(
  response=predict(fit_prev1,newdata=b.dataPb,type='response'),
  variance=predict(fit_prev1,newdata=b.dataPb,type='variance'),
  predict(fit_prev1,newdata=b.dataPb,type='quantile',at=c(0.025,0.975)))
prevb <- as.data.frame(predPb)
prevb <- cbind(prevb,b.dataPb)
b <- ggplot(prevb,aes(x=BladeAreaMean))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=BladeAreaMean,y=PrevalenceMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab(expression("Blade Area (cm"^"2"~")"))+
  ylab("Wasting disease prevalence\n (% individuals infected)")+
  theme_bw()+
  theme(panel.grid = element_blank())
a/b + plot_layout(guides="collect")
```

Finally, calculate standardized coefficients (can do this by scaling the inputs to the model?)
```{r prev_model_stdcoef}
SDy <- sd(dat$PrevalenceMean)
SDx1 <- sd(dat$CPTempAnomaly)
SDx2 <- sd(dat$BladeAreaMean)
coefficients(fit_prev1)[2]*SDx1
coefficients(fit_prev1)[3]*SDx2
# note, can also get these coeffs by scaling the predictors prior to fitting the model
# dat$sCPTA <- scale(dat$CPTempAnomaly,scale=TRUE,center=TRUE)
# dat$sBA <- scale(dat$BladeAreaMean,scale=TRUE,center=TRUE)
# fit_prev1s <- betareg(PrevalenceMean~sCPTA + sBA,
#                      data=dat,
#                      weights = Count,
#                      link = "logit")
# summary(fit_prev1s)
```
Standardized coefficient for CPTA is about 5x greater than the standardized coefficient for blade area.

### Severity model

Repeat same analysis for severity

```{r sev_model_build, echo=FALSE}
fit_sev1 <- betareg(SeverityMean~CPTempAnomaly + BladeAreaMean,
                     data=dat,
                     weights = Count,
                     link = "logit")

fit_sev2 <- betareg(SeverityMean~CPTempAnomaly + DensityShootsMean,
                     data=dat,
                     weights = Count,
                     link = "logit")
fit_sev3 <- betareg(SeverityMean~CPTempAnomaly,
                     data=dat,
                     weights = Count,
                     link = "logit")
fit_sev4 <- betareg(SeverityMean~ BladeAreaMean,
                     data=dat,
                     weights = Count,
                     link = "logit")
fit_sev5 <- betareg(SeverityMean~ DensityShootsMean,
                     data=dat,
                     weights = Count,
                     link = "logit")
df.AIC <- AIC(fit_sev1,fit_sev2,fit_sev3,fit_sev4,fit_sev5)
df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
Note, the model with three fixed predictors is better by AIC for severity. BUT this model is overfitted - the p-values are suddenly extremely small, including for CTempAnomaly. Which doesn't make sense. So I think we cannot have three predictors on this small of a dataset. 

Second note, the models with shoot density are worse by AIC than models with blade area. Although these are different predictors (not nested) but I believe we can still do AIC, just not a log-likelihood.

Use the same model as for prevalence (want to know if the same factors drive both).

```{r sev_model_eval,echo=FALSE,fig.width=4,fig.height=3}
summary(fit_sev1)
E.sev <- resid(fit_sev1,type='pearson')
F.sev <- fitted(fit_sev1)
plot(E.sev~F.sev)
plot(E.sev~dat$BladeAreaMean)
plot(E.sev~dat$CPTempAnomaly)

## Visualize the best-fitting model effect sizes
sev_names <- c("Leaf area", "Cumulative SST anomaly")
meadow_plot2 <- plot_model(fit_sev1,
                          type="std",
           axis.labels = sev_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.2,
           #axis.lim = c(0.2,5),
           group.terms = c(1,2)
           )
meadow_plot2+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.5,2))+
  scale_color_manual(values=c("grey50", "black"))+
  ylab("Scaled estimates of \ndisease severity ratio")+
  #xlab("Scaled parameters")+
  labs(tag="B")+
  theme(panel.grid = element_blank(),
        axis.text = element_text(size=10))
```

Beach Haven residual is a bit of an outlier perhaps but otherwise no major issues with the residuals when plotted against covariates and fitted values. Model is acceptable.

Blade area is significant, CPTA is not. Visualize model by simulating fit across new data holding one variable constant at a time.

```{r sev_model_viz,echo=FALSE,fig.width=5,fig.height=7}
jvaluesSc <- with(dat, seq(from = min(BladeAreaMean), to = max(BladeAreaMean), length.out = 100))
b.dataSc <- data.frame(BladeAreaMean=jvaluesSc,CPTempAnomaly=median(dat$CPTempAnomaly),Region="BC")
predSc <- cbind(
  response=predict(fit_sev1,newdata=b.dataSc,type='response'),
  variance=predict(fit_sev1,newdata=b.dataSc,type='variance'),
  predict(fit_sev1,newdata=b.dataSc,type='quantile',at=c(0.025,0.975)))
sevc <- as.data.frame(predSc)
sevc <- cbind(sevc,b.dataSc)
c <- ggplot(sevc,aes(x=BladeAreaMean))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=BladeAreaMean,y=SeverityMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab(expression("Blade Area (cm"^"2"~")"))+
  ylab("Wasting disease severity\n (% blade area damaged)")+
  theme_bw()+
  theme(panel.grid = element_blank())

# Don't simulate with CPTA because it's not significant
# jvaluesS <- with(dat, seq(from = min(CPTempAnomaly), to = max(CPTempAnomaly), length.out = 100))
# median(dat$CPTempAnomaly)
# mean(dat$CPTempAnomaly)
# b.dataS <- data.frame(CPTempAnomaly=jvaluesS,BladeAreaMean=median(dat$BladeAreaMean),Region="BC")
# predS <- cbind(
#   response=predict(fit_sev1,newdata=b.dataS,type='response'),
#   variance=predict(fit_sev1,newdata=b.dataS,type='variance'),
#   predict(fit_sev1,newdata=b.dataS,type='quantile',at=c(0.025,0.975)))
# sev <- as.data.frame(predS)
# sev <- cbind(sev,b.dataS)
d <- ggplot()+
  #geom_line(aes(y=response))+
  #geom_line(aes(y=q_0.025),linetype="dashed")+
  #geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=CPTempAnomaly,y=SeverityMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab("Cumulative positive temperature anomaly (ºC)")+
  ylab("Wasting disease severity\n (% blade area damaged)")+
  theme_bw()+
  theme(panel.grid = element_blank())
(d / c) + plot_layout(guides="collect")
```


```{r sev_model_stdcoef}
SDy <- sd(dat$PrevalenceMean)
SDx1 <- sd(dat$CPTempAnomaly)
SDx2 <- sd(dat$BladeAreaMean)
coefficients(fit_sev1)[2]*SDx1
coefficients(fit_sev1)[3]*SDx2
# note, can also get these coeffs by scaling the predictors prior to fitting the model
# dat$sCPTA <- scale(dat$CPTempAnomaly,scale=TRUE,center=TRUE)
# dat$sBA <- scale(dat$BladeAreaMean,scale=TRUE,center=TRUE)
# fit_sev1s <- betareg(SeverityMean~sCPTA + sBA,
#                      data=dat,
#                      weights = Count,
#                      link = "logit")
# summary(fit_sev1s)
```

Here, the standardized coeff for blade area is 30x greater in magnitude than the CPTA (which is non-sig anyway).

Overall, CPTA was significant for determining prevalence but not for severity, which is significantly affected by blade area.

### Figure for paper

Making a nice looking figure for the paper woo

```{r figure_paper,fig.width=4.75,fig.height=5.5}
a1 <- ggplot(preva,aes(x=CPTempAnomaly))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=CPTempAnomaly,y=PrevalenceMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  guides(color = guide_legend(nrow = 1))+
  xlab("Cumulative positive \ntemperature anomaly (ºC)")+
  ylab("Wasting disease prevalence\n (% individuals infected)")+
  theme_bw(base_size = 9)+
  theme(panel.grid = element_blank(),
        legend.margin = margin(t=0,r=0,b=0,l=1,unit="pt"),
        legend.key.size = unit(5,unit="mm"),
        legend.text = element_text(size=8),
        legend.title = element_blank(),
        legend.position = "bottom",
        legend.justification = "center",
        legend.direction = "horizontal"
        )
b1 <-ggplot(prevb,aes(x=BladeAreaMean))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=BladeAreaMean,y=PrevalenceMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab(expression("Blade Area (cm"^"2"~")"))+
  ylab("Wasting disease prevalence\n (% individuals infected)")+
  theme_bw(base_size = 9)+
  theme(panel.grid = element_blank(),
        legend.margin = margin(t=0,r=0,b=0,l=1,unit="pt"),
        legend.key.size = unit(5,unit="mm"),
        legend.title = element_blank(),
        legend.position = "")
c1 <- ggplot(sevc,aes(x=BladeAreaMean))+
  geom_line(aes(y=response))+
  geom_line(aes(y=q_0.025),linetype="dashed")+
  geom_line(aes(y=q_0.975),linetype="dashed")+
  geom_point(data=dat,aes(x=BladeAreaMean,y=SeverityMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab(expression("Blade Area (cm"^"2"~")"))+
  ylab("Wasting disease severity\n (% blade area damaged)")+
  theme_bw(base_size = 9)+
  theme(panel.grid = element_blank(),
        legend.margin = margin(t=0,r=0,b=0,l=1,unit="pt"),
        legend.key.size = unit(5,unit="mm"),
        legend.title = element_blank(),
        legend.position = "")
d1 <- ggplot()+
  geom_point(data=dat,aes(x=CPTempAnomaly,y=SeverityMean,color=Region),size=2)+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+ 
  scale_color_viridis_d()+
  xlab("Cumulative positive \ntemperature anomaly (ºC)")+
  ylab("Wasting disease severity\n (% blade area damaged)")+
  theme_bw(base_size = 9)+
  theme(panel.grid = element_blank(),
        legend.margin = margin(t=0,r=0,b=0,l=1,unit="pt"),
        legend.key.size = unit(5,unit="mm"),
        legend.title = element_blank(),
        legend.position = "")

#pcombo <- a1 / b1
#scombo <- d1 / c1 
#(pcombo | scombo) + plot_layout(guides="collect") + plot_annotation(tag_levels = "A")

## trying again with cowplot
a1a <- a1 +theme(legend.position = "")
legend <- get_legend(a1+theme(legend.box.margin = margin(6,0,0,0),
                              legend.direction = "horizontal"))
pcombo <- cowplot::plot_grid(a1a,b1,nrow=2, labels=c("A","B"))
scombo <- cowplot::plot_grid(d1,c1,nrow=2, labels=c("C","D"))
total <- cowplot::plot_grid(pcombo,scombo,ncol=2)
total_l <- cowplot::plot_grid(total,legend,nrow=2,rel_heights = c(1,.1))
total_l
#plot_grid(a1,b1,c1,d1,nrow = 2,legend,rel_heights = c(1,.1))
```