---
title: "Transect_glmm"
author: "LRA"
date: "1/26/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(patchwork)
library(performance)
library(sjPlot)
region_order <- c("AK","BC","WA","OR","BB","SD")

```

## Transect level models of disease

Two sets of models, one with cumulative positive temperature anomaly as a predictor, one without.

Best models are selected based on AIC and model weight within the model set.

Because only 27 meadows out of 32 have SST data for the anomaly calculation, the number of replicates is smaller for those models.

However, the significant effects are mainly the same across the models.

Predictors are: Blade Area, Shoot Density, Epiphyte Mass per Blade Area, Cumulative positive anomaly, Tidal Height, and all interactions between Tidal Height and fixed effects.

Random structure for transects is Meadow within Region. For the restircted dataset (SST sites only), Region as a random effect causes a singular model (not enough variaiton in Region, also the design is highly unbalanced). Therefore for those models, Region is not included. 

Note, for severity, transect-level severity is summarized based on diseased blades only. Predictor values are the same.

```{r data}
transect <- read.csv("Refined Models/input/transect_combined_parameters.csv")
transect$Region <- ordered(transect$Region,levels=region_order)
transect$Meadow <- as.factor(transect$Meadow)
transect$sBladeAreaMean <- scale(transect$BladeAreaMean,scale=TRUE,center=TRUE)
transect$sDensityShootsMean <- scale(transect$DensityShootsMean,scale=TRUE,center=TRUE)
transect$sEpiphytePerAreaMean <- scale(transect$EpiphytePerAreaMean,scale=TRUE,center=TRUE)
transect$sCDiffMeanHeat <- scale(transect$CDiffMeanHeat,scale=TRUE,center=TRUE)
dat <- transect[-which(is.na((transect$CDiffMeanHeat))),]
dat$Region <- ordered(dat$Region,levels=region_order)
dat$sBladeAreaMean <- scale(dat$BladeAreaMean,scale=TRUE,center=TRUE)
dat$sDensityShootsMean <- scale(dat$DensityShootsMean,scale=TRUE,center=TRUE)
dat$sEpiphytePerAreaMean <- scale(dat$EpiphytePerAreaMean,scale=TRUE,center=TRUE)
dat$sCDiffMeanHeat <- scale(dat$CDiffMeanHeat,scale=TRUE,center=TRUE)

# data exploration
# transect_dat <- select(transect,c("TidalHeight","LongestBladeLengthMean","EpiphytePerAreaMean","CDiffMeanHeat",
#                                   "DensityShootsMean","BladeAreaMean","PrevalenceMean","SeverityMean"))
# transect_dat$TidalHeight <- as.factor(transect_dat$TidalHeight)
# pairs(transect_dat)
```

## Prevalence model with CPTA n=162
predictors: Density, Blade Area, Epiphyte mass per area, CPTA, Tidal height, and interactions between fixed effects and TH
random structure: site within region

```{r transect_prev_cpta_selection, echo=FALSE}
fit_prev1s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                      +Region+(1|Meadow),
                    data=dat,
                    weights=CountBlades,
                    family=binomial)

# fixed effects are scaled for easier comparison and to avoid convergence issues
# Note, Region as a random effect (1|Region) results in a singular model (variance of Region is 0)
# This suggests the variation between regions is better explained in the residual than as a random effect (or maybe there are
# not enough regions, although 6 should be sufficient)
# fit_prev1s includes Region as a fixed effect
# drop1(fit_prev1s)
# # by AIC, dropping Region as a fixed effect improves the model by >2, therefore drop Region
fit_prev2s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)

 ### Model selection via AIC, using a priori model set
 # remove blade area : tidal height interaction
 fit_prev2.1s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
 
 # remove shoot density : tidal height interaction
fit_prev2.2s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
  # remove epiphyte load : tidal height interaction
fit_prev2.3s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
 # remove cumulative anomaly : tidal height interaction
 fit_prev2.4s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
 # no interactions
 fit_prev2.5s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 # no tidal height (or interactions)
 fit_prev2.6s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)

 df.AIC <- data.frame(AIC(fit_prev2s,fit_prev2.1s,fit_prev2.2s,fit_prev2.3s,fit_prev2.4s,
     fit_prev2.5s,fit_prev2.6s))
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
# Based on AIC and model weights, best model is fit_prev2.1s, second model (no blade area:tidal height interaction)
```
Based on AIC and model weights, best model is fit_prev2.1s, second model (no blade area:tidal height interaction)

Now need to validate the model and record output

```{r transect_prev_cpta_output, echo=FALSE,fig.width=4,fig.height=4}
 
# Assess the model residuals 
E2.sim <- simulateResiduals(fit_prev2.1s)
plot(E2.sim)
plot(E2.sim$scaledResiduals~dat$sBladeAreaMean)
plot(E2.sim$scaledResiduals~dat$sDensityShootsMean)
plot(E2.sim$scaledResiduals~dat$sCDiffMeanHeat)
plot(E2.sim$scaledResiduals~as.factor(dat$TidalHeight))

# Look at the model output, performance metrics
summary(fit_prev2.1s)
performance(fit_prev2.1s)

## Visualize the best-fitting model output
p1names <- c("Cumulative SST anomaly * Tidal height (upper)", "Epiphyte load * Tidal height (upper)",
             "Shoot density * Tidal height (upper)",
             "Tidal height (upper)","Cumulative SST anomaly","Epiphyte load","Shoot density","Leaf area")
p1 <- plot_model(fit_prev2.1s,
           axis.labels = p1names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.3,
           axis.lim = c(0.2,5),
           group.terms = c(1,2,1,2,1,2,1,2)
           )
p1+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("grey50","black"))+
  ylab("Scaled estimates of \n disease prevalence odds ratio")+
  #xlab("Scaled parameters")+
  labs(tag="A")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))

```

Some issues with the uniformity of the simulated residuals, but there's not significant dispersion so it is probably okay to use the model. 
 
Scaled odds ratio estimates indicate the change in odds of disease prevalence, given a change of 1 SD of the predictor. E.g. for a given transect, an increase of 1 SD of the cumulative SST anomaly means the transect is 2.46 times more likely to be diseased. 
 
##  Transect-level prevalence model, for all transects, no CPTA, n=192
```{r transect_prev_all_selection}
transect$Meadow <- paste(transect$Region,transect$SiteCode,sep="_")
fit_prev5 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)

# summary(fit_prev5)
# no issue with Region as a random effect for the full dataset (more balanced?)

### Model selection by AIC
 # remove blade area : tidal height interaction
 fit_prev5.1 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 
 # remove shoot density : tidal height interaction
fit_prev5.2 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
  # remove epiphyte load : tidal height interaction
fit_prev5.3 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no interactions
 fit_prev5.4 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no tidal height (or interactions)
 fit_prev5.5 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 
 df.AIC <- AIC(fit_prev5,fit_prev5.1,fit_prev5.2,fit_prev5.3,fit_prev5.4,fit_prev5.5)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
By AIC and weight, model 5.1 is the best-fitting (no blade area interaction with tidal height).

```{r transect_prev_all_output, echo=FALSE,fig.width=4,fig.height=4}
 
# Assess the model residuals 
E5.sim <- simulateResiduals(fit_prev5.1)
plot(E5.sim)
plot(E5.sim$scaledResiduals~transect$sBladeAreaMean)
plot(E5.sim$scaledResiduals~transect$sDensityShootsMean)
plot(E5.sim$scaledResiduals~as.factor(transect$TidalHeight))

# Look at the model output, performance metrics
summary(fit_prev5.1)
performance(fit_prev5.1)
## Visualize the best-fitting model output
p2names <- c("Epiphyte load * Tidal height (upper)","Shoot density * Tidal height (upper)",
             "Tidal height (upper)","Epiphyte load","Shoot density","Leaf area")
p2 <- plot_model(fit_prev5.1,
           axis.labels = p2names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.3,
           axis.lim = c(0.2,10),
           group.terms = c(1,1,1,1,1,1)
           )
p2+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,4))+
  scale_color_manual(values=c("black"))+
  ylab("Scaled estimates of \ndisease prevalence odds ratio")+
  labs(tag = "C")+
  #xlab("Scaled parameters")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))

```
Scaled odds ratio estimates indicate the change in odds of disease prevalence, given a change of 1 SD of the predictor. E.g. for a given transect, an increase of 1 SD of the shoot density means the transect is 1.7 times more likely to be diseased.

Same parameter estimates for the models with and without CPTA (restricted vs not restricted), but restricted model has a higher marginal R2. Therefore, temperature is perhaps important for explaining the prevalence dynamics.

## Transect level severity with CPTA n = 155
```{r transect_sev_cpta_selection}
# transect-level severity means are calculated using hurdle approach (only diseased blades are included)
# nine transects had zero diseaed blades, so exclude those nine for the base severity dataset
sevT <- subset(transect,SeverityMean>0)
sevTdat <- sevT[-which(is.na((sevT$CDiffMeanHeat))),]

## fit with CPTA to start (data = sevTdat)
# fit_sev1 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
#                      sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                       (1|Region)+(1|Meadow),
#                     data=sevTdat,
#                     family=beta_family(link = "logit"))
# summary(fit_sev1)
# # again Region variance is 0 because of unevenness, can try fitting Region as a fixed effect but this increases the AIC
#  fit_sev2 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
#                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                        Region+(1|Meadow),
#                      data=sevTdat,
#                      family=beta_family(link = "logit"))
#  summary(fit_sev2)
# drop1(fit_sev2)
# drop Region from the model
fit_sev3 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|Meadow),
                    data=sevTdat,
                    family=beta_family(link = "logit"))
# summary(fit_sev3)
# Model selection via AIC
 # remove blade area : tidal height interaction
 fit_sev3.1s <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sDensityShootsMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|Meadow),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
 
 # remove shoot density : tidal height interaction
fit_sev3.2s <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeAreaMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|Meadow),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
  # remove epiphyte load : tidal height interaction
fit_sev3.3s <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       sCDiffMeanHeat:TidalHeight+
                      (1|Meadow),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
 # remove cumulative anomaly : tidal height interaction
 fit_sev3.4s <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+
                      (1|Meadow),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
 # no interactions
 fit_sev3.5s <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       (1|Meadow),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 # no tidal height (or interactions)
 fit_sev3.6s <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+
                      (1|Meadow),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 df.AIC <- AIC(fit_sev3,fit_sev3.1s,fit_sev3.2s,fit_sev3.3s,fit_sev3.4s,fit_sev3.5s,fit_sev3.6s)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
By AIC and weight, the best model is fit_sev3.6 (no tidal height and no interactions)

```{r transect_sev_cpta_output,fig.width=4,fig.height=4}
# Assess the model residuals 
E3.sim <- simulateResiduals(fit_sev3.6s)
plot(E3.sim)
plot(E3.sim$scaledResiduals~sevTdat$sBladeAreaMean)
plot(E3.sim$scaledResiduals~sevTdat$sDensityShootsMean)
plot(E3.sim$scaledResiduals~sevTdat$sEpiphytePerAreaMean)
plot(E3.sim$scaledResiduals~as.factor(sevTdat$TidalHeight))
plot(E3.sim$scaledResiduals~sevTdat$sCDiffMeanHeat)

# Look at the model output, performance metrics
summary(fit_sev3.6s)
performance(fit_sev3.6s)
## Visualize the best-fitting model output
s_cpta_names <- c("Cumulative SST anomaly","Epiphyte load","Shoot density","Leaf area")
s_cpta_plot <- plot_model(fit_sev3.6s,
           axis.labels = s_cpta_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.3,
            #transform = NULL
           #axis.lim = c(0.1,10),
           
           group.terms = c(1,2,2,2)
           )
s_cpta_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("black","grey50"))+
  ylab("Scaled estimates of \ndisease severity odds ratio")+
  labs(tag="B")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))

```
In contrast to prevalence models, severity model is beta regression, so the scaled parameter estimates estimate the change in the ratio of proportions, rather than the ratio of probabilities (odds). E.g. for an increase in leaf area of 1 SD, the ratio of diseased to non-diseased tissue will change by a factor of 0.70. Only leaf area is significant; effect sizes are smaller and non-significant for other predictors, including cumulative SST anomaly. 

## Transect level severity without CPTA n = 183 ##### 
```{r transect_sev_all_selection}
# fit models using sevT, all transects, no CPTA 
# fit_sev1 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
#                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
#                        sEpiphytePerAreaMean:TidalHeight+
#                       (1|Region)+(1|Meadow),
#                     data=sevT,
#                     family=beta_family(link = "logit"))
# # summary(fit_sev1)
# # Region is also close to zero, try as fixed effect but it increases AIC
# fit_sev2 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
#                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
#                        sEpiphytePerAreaMean:TidalHeight+
#                       Region+(1|Meadow),
#                     data=sevT,
#                     family=beta_family(link = "logit"))
# drop1(fit_sev2)
# drop Region
# basic model for selection
fit_sev4 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+
                      (1|Meadow),
                    data=sevT,
                    family=beta_family(link = "logit"))

 # Model selection
 # remove blade area : tidal height interaction
 fit_sev4.1 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sDensityShootsMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+
                      (1|Meadow),
                    data=sevT,
                    family=beta_family(link = "logit"))
 
 # remove shoot density : tidal height interaction
fit_sev4.2 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+
                      (1|Meadow),
                    data=sevT,
                    family=beta_family(link = "logit"))
  # remove epiphyte load : tidal height interaction
fit_sev4.3 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       (1|Meadow),
                    data=sevT,
                    family=beta_family(link = "logit"))
 # no interactions
 fit_sev4.4 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                      (1|Meadow),
                    data=sevT,
                    family=beta_family(link = "logit"))
 
 # no tidal height (or interactions)
 fit_sev4.5 <- glmmTMB(SeverityMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+
                      (1|Meadow),
                    data=sevT,
                    family=beta_family(link = "logit"))
 df.AIC <- AIC(fit_sev4,fit_sev4.1,fit_sev4.2,fit_sev4.3,fit_sev4.4,
     fit_sev4.5)
 df.AIC$deltaAIC <- df.AIC$AIC-min(df.AIC$AIC)
 df.AIC$likelihood <- exp(-df.AIC$deltaAIC/2)
 df.AIC$weight <- df.AIC$likelihood/sum(df.AIC$likelihood)
 df.AIC
```
By AIC and weight, best model is fit_sev4.1, no blade area: tidal height interaction

```{r transect_sev_all_output,fig.width=4,fig.height=4}
# Assess the model residuals 
E4.sim <- simulateResiduals(fit_sev4.1)
plot(E4.sim)
plot(E4.sim$scaledResiduals~sevT$sBladeAreaMean)
plot(E4.sim$scaledResiduals~sevT$sDensityShootsMean)
plot(E4.sim$scaledResiduals~sevT$sEpiphytePerAreaMean)
plot(E4.sim$scaledResiduals~as.factor(sevT$TidalHeight))

# Look at the model output, performance metrics
summary(fit_sev4.1)
performance(fit_sev4.1)

## Visualize the best-fitting model output
s_all_names <- c("Epiphyte load * Tidal height (upper)","Shoot density * Tidal height (upper)",
                "Tidal height (upper)","Epiphyte load","Shoot density","Leaf area")
s_all_plot <- plot_model(fit_sev4.1,
           axis.labels = s_all_names,
           title="",
           show.p = TRUE,
           show.values = TRUE,
           value.offset = 0.3,
            #transform = NULL
           #axis.lim = c(0.1,10),
           
           group.terms = c(1,1,2,2,1,2)
           )
s_all_plot+theme_bw()+
  geom_hline(yintercept = 1,linetype="dashed",color="darkgrey")+
  scale_y_log10(limits=c(0.2,5))+
  scale_color_manual(values=c("black","grey50"))+
  ylab("Scaled estimates of \ndisease severity odds ratio")+
  labs(tag="D")+
  theme(panel.grid = element_blank(),
        #plot.margin = margin(0,0,0,0,unit="pt"),
        axis.text = element_text(size=10))
```

Scaled estimates indicate change in the ratio of diseased tissue to non-diseased tissue. For a 1 SD increase in shoot density, the ratio of diseased tissue to non-diseased tissue changes by a factor of 1.84. 

Same estimate for leaf area but for the unrestricted dataset (all transects), shoot density is significant. 
