---
title: "Transect_glmm"
author: "LRA"
date: "1/26/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(TMB)
library(glmmTMB)
library(DHARMa)
library(patchwork)
library(performance)
region_order <- c("AK","BC","WA","OR","BB","SD")

```

## Transect level models of disease

Two sets of models, one with cumulative positive temperature anomaly as a predictor, one without.

Because only 27 meadows out of 32 have SST data for the anomaly calculation, the number of replicates is smaller for those models.

However, the significant effects are mainly the same across the models.

Predictors are: Blade Area, Shoot Density, Epiphyte Mass per Blade Area, Cumulative positive anomaly, Tidal Height, and all interactions between Tidal Height and fixed effects.

Random structure for transects is Meadow within Region.

Note, for severity, transect-level severity is re-summarized based on diseased blades only. Predictor values are the same.

```{r data}
transect <- read.csv("Refined Models/input/transect_combined_parameters.csv")
transect$Region <- ordered(transect$Region,levels=region_order)
transect$Meadow <- as.factor(transect$Meadow)
transect$sBladeAreaMean <- scale(transect$BladeAreaMean,scale=TRUE,center=TRUE)
transect$sDensityShootsMean <- scale(transect$DensityShootsMean,scale=TRUE,center=TRUE)
transect$sEpiphytePerAreaMean <- scale(transect$EpiphytePerAreaMean,scale=TRUE,center=TRUE)
transect$sCDiffMeanHeat <- scale(transect$CDiffMeanHeat,scale=TRUE,center=TRUE)
dat <- transect[-which(is.na((transect$CDiffMeanHeat))),]
dat$Region <- ordered(dat$Region,levels=region_order)
dat$sBladeAreaMean <- scale(dat$BladeAreaMean,scale=TRUE,center=TRUE)
dat$sDensityShootsMean <- scale(dat$DensityShootsMean,scale=TRUE,center=TRUE)
dat$sEpiphytePerAreaMean <- scale(dat$EpiphytePerAreaMean,scale=TRUE,center=TRUE)
dat$sCDiffMeanHeat <- scale(dat$CDiffMeanHeat,scale=TRUE,center=TRUE)

# data exploration
# transect_dat <- select(transect,c("TidalHeight","LongestBladeLengthMean","EpiphytePerAreaMean","CDiffMeanHeat",
#                                   "DensityShootsMean","BladeAreaMean","PrevalenceMean","SeverityMean"))
# transect_dat$TidalHeight <- as.factor(transect_dat$TidalHeight)
# pairs(transect_dat)
```

## Prevalence model with CPTA n=162
predictors: Density, Blade Area, Epiphyte mass per area, CPTA, Tidal height, and interactions between fixed effects and TH
random structure: site within region

```{r transect_prev_cpta, echo=FALSE}
fit_prev1s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                      +Region+(1|Meadow),
                    data=dat,
                    weights=CountBlades,
                    family=binomial)
# scaling removes the issue with convergence
# Note, Region as a random effect (1|Region) results in a singluar model (variance of Region is 0)
# This suggests the variation between regions is better explained in the residual than as a random effect (or maybe there are
# not enough regions, although 6 should be sufficient)
# fit_prev1s includes Region as a fixed effect
summary(fit_prev1s)
performance(fit_prev1s)
icc(fit_prev1s,by_group = FALSE)
class(transect$Region)
E1.sim <- simulateResiduals(fit_prev1s)
plot(E1.sim)
plot(E1.sim$scaledResiduals~dat$sBladeAreaMean)
plot(E1.sim$scaledResiduals~dat$sDensityShootsMean)
plot(E1.sim$scaledResiduals~as.factor(dat$TidalHeight))
plot(E1.sim$scaledResiduals~dat$Region)
drop1(fit_prev1s)
# # by AIC, dropping Region as a fixed effect improves the model by >2, therefore drop Region
fit_prev2s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial,dispformula = ~Region)
performance(fit_prev2s)
E2.sim <- simulateResiduals(fit_prev2s)
plot(E2.sim)
plot(E2.sim$scaledResiduals~dat$sBladeAreaMean)
plot(E2.sim$scaledResiduals~dat$sDensityShootsMean)
plot(E2.sim$scaledResiduals~as.factor(dat$TidalHeight))
plot(E2.sim$scaledResiduals~dat$Region)
drop1(fit_prev2s)
# summary(fit_prev2s)
#drop sBladeAreaMean:TidalHeight
fit_prev3s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
summary(fit_prev3s)
performance(fit_prev3s)
icc(fit_prev3s,by_group=TRUE)
# anova(fit_prev2s,fit_prev3s)
# # no difference, so can use simpler model
 drop1(fit_prev3s)
# # no other terms warrant being dropped
# # but, how do we interpret this? The slope of scaled shoot density varies by tidal height, but shoot density and tidal height
# # are not significant on their own
# # try dropping this interaction
 fit_prev4s <- glmmTMB(PrevalenceMean~sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 anova(fit_prev3s,fit_prev4s)
drop1(fit_prev4s)
performance(fit_prev4s) 
# # gives a significant effect, so need to keep the interaction
# # so keep fit_prev3s as the final model?
 E3.sim <- simulateResiduals(fit_prev3s)
 plot(E3.sim)
 plot(E3.sim$scaledResiduals~dat$sBladeAreaMean)
 plot(E3.sim$scaledResiduals~dat$sDensityShootsMean)
 plot(E3.sim$scaledResiduals~as.factor(dat$TidalHeight))
 plot(E3.sim$scaledResiduals~dat$Region)
 # residuals are acceptable, keep the model
 
 ### different way to do this is to look at AIC of specific models
 # remove blade area : tidal height interaction
 fit_prev2.1s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
 
 # remove shoot density : tidal height interaction
fit_prev2.2s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
  # remove epiphyte load : tidal height interaction
fit_prev2.3s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
 # remove cumulative anomaly : tidal height interaction
 fit_prev2.4s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+
                        (1|Meadow),
                      data=dat,
                      weights=CountBlades,
                      family=binomial)
 # no interactions
 fit_prev3.1s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 # no blade area
 fit_prev3.2s <- glmmTMB(PrevalenceMean~sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 # no shoot density
 fit_prev3.3s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 # no epiphyte load
 fit_prev3.4s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 # no cumulative anomaly
 fit_prev3.5s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 # no tidal height (or interactions)
 fit_prev3.6s <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+
                       (1|Meadow),
                     data=dat,
                     weights=CountBlades,
                     family=binomial)
 AIC(fit_prev2s,fit_prev2.1s,fit_prev2.2s,fit_prev2.3s,fit_prev2.4s,
     fit_prev3.1s,fit_prev3.2s,fit_prev3.3s,fit_prev3.4s,fit_prev3.5s,fit_prev3.6s)
```

 
##  Transect-level prevalence model, for all transects, no CPTA, n=192
```{r transect_prev_all}
transect$Meadow <- paste(transect$Region,transect$SiteCode,sep="_")
fit_prev5 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)

summary(fit_prev5)
performance(fit_prev5)
icc(fit_prev5,by_group = TRUE)
drop1(fit_prev5)

### different way to do this is to look at AIC of specific models
 # remove blade area : tidal height interaction
 fit_prev5.1 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 
 # remove shoot density : tidal height interaction
fit_prev5.2 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
  # remove epiphyte load : tidal height interaction
fit_prev5.3 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no interactions
 fit_prev5.4 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no blade area
 fit_prev5.5 <- glmmTMB(PrevalenceMean~sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                      sDensityShootsMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no shoot density
 fit_prev5.6 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sEpiphytePerAreaMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no epiphyte load
 fit_prev5.7 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+TidalHeight+
                       sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 # no tidal height (or interactions)
 fit_prev5.8 <- glmmTMB(PrevalenceMean~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+
                       +(1|Region)+(1|Meadow),
                     data=transect,
                     weights=CountBlades,
                     family=binomial)
 AIC(fit_prev5,fit_prev5.1,fit_prev5.2,fit_prev5.3,fit_prev5.4,
     fit_prev5.5,fit_prev5.6,fit_prev5.7,fit_prev5.8)
```

## transect level severity with CPTA n = 155 ####
```{r transect_sev_cpta}
sev <- read.csv("Refined Models/input/severity_blades.csv")
sevT <- sev %>%
  group_by(Region,SiteCode,TidalHeight,Transect)%>%
  summarise(across(c("Severity","BladeArea","DensityShoots","EpiphytePerAreaMean","CDiffMeanHeat",
                     "sBladeArea","sDensityShoots","sEpiphytePerAreaMean","sCDiffMeanHeat"),mean))
sevT$MeadowId <- paste(sevT$Region,sevT$SiteCode,sep="_")
sevT$TransectId <- paste(sevT$MeadowId,sevT$Transect,sep="_")

sevTdat <- sevT[-which(is.na((sevT$CDiffMeanHeat))),]
sevTdat$sBladeAreaMean <- scale(sevTdat$BladeArea,scale=TRUE,center=TRUE)
sevTdat$sDensityShootsMean <- scale(sevTdat$DensityShoots,scale=TRUE,center=TRUE)
sevTdat$sEpiphytePerAreaMean <- scale(sevTdat$EpiphytePerAreaMean,scale=TRUE,center=TRUE)
sevTdat$sCDiffMeanHeat <- scale(sevTdat$CDiffMeanHeat,scale=TRUE,center=TRUE)

## fit with CPTA to start (data = sevTdat)
# fit_sev1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
#                       (1|Region)+(1|MeadowId),
#                     data=sevTdat,
#                     family=beta_family(link = "logit"))
# summary(fit_sev1)
# # again region variance is wonky because of unevenness
 fit_sev2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                       sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       Region+(1|MeadowId),
                     data=sevTdat,
                     family=beta_family(link = "logit"))
# summary(fit_sev2)
# drop1(fit_sev2)
# drop Region
fit_sev3 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                      sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                      (1|MeadowId),
                    data=sevTdat,
                    family=beta_family(link = "logit"))
summary(fit_sev3)
# drop1(fit_sev3)
# fit_sev4 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
#                       (1|MeadowId),
#                     data=sevTdat,
#                     family=beta_family(link = "logit"))
# anova(fit_sev4,fit_sev3)
# # no difference by AIC but log-likelihood test suggests we don't need it 
# fit_sev5 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
#                       (1|MeadowId),
#                     data=sevTdat,
#                     family=beta_family(link = "logit"))
# anova(fit_sev5,fit_sev4)
# also can drop the interactions?
# fit_sev6 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
#                       sDensityShoots:TidalHeight+
#                       (1|MeadowId),
#                     data=sevTdat,
#                     family=beta_family(link = "logit"))
# anova(fit_sev6,fit_sev5)
# drop1(fit_sev6)
# summary(fit_sev6)
# summary(fit_sev3)
# honestly I don't think it matters which to use... a model selection preference
# go with fit_sev3
performance(fit_sev3)
icc(fit_sev3,by_group = TRUE)
# this seems wrong (plus the warning message..) but maybe report just the total ICC?
#check residuals on fit_sev3
# Es.sim <- simulateResiduals(fit_sev3)
# plot(Es.sim)
# plot(Es.sim$scaledResiduals~sevTdat$sBladeArea)
# plot(Es.sim$scaledResiduals~sevTdat$sDensityShoots)
# plot(Es.sim$scaledResiduals~as.factor(sevTdat$TidalHeight))
# plot(Es.sim$scaledResiduals~sevTdat$Region)

 ### different way to do this is to look at AIC of specific models
 # remove blade area : tidal height interaction
 fit_sev2.1s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|MeadowId),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
 
 # remove shoot density : tidal height interaction
fit_sev2.2s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                        (1|MeadowId),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
  # remove epiphyte load : tidal height interaction
fit_sev2.3s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                        (1|MeadowId),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
 # remove cumulative anomaly : tidal height interaction
 fit_sev2.4s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+
                        (1|MeadowId),
                      data=sevTdat,
                    family=beta_family(link = "logit"))
 # no interactions
 fit_sev3.1s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                       (1|MeadowId),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 # no blade area
 fit_sev3.2s <- glmmTMB(Severity~sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 # no shoot density
 fit_sev3.3s <- glmmTMB(Severity~sBladeAreaMean+sEpiphytePerAreaMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 # no epiphyte load
 fit_sev3.4s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sCDiffMeanHeat+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sCDiffMeanHeat:TidalHeight+
                       (1|MeadowId),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 # no cumulative anomaly
 fit_sev3.5s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+TidalHeight+
                        sBladeAreaMean:TidalHeight+sDensityShootsMean:TidalHeight+
                        sEpiphytePerAreaMean:TidalHeight+
                       (1|MeadowId),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 # no tidal height (or interactions)
 fit_sev3.6s <- glmmTMB(Severity~sBladeAreaMean+sDensityShootsMean+sEpiphytePerAreaMean+sCDiffMeanHeat+
                       (1|MeadowId),
                     data=sevTdat,
                    family=beta_family(link = "logit"))
 AIC(fit_sev3,fit_sev2.1s,fit_sev2.2s,fit_sev2.3s,fit_sev2.4s,
     fit_sev3.1s,fit_sev3.2s,fit_sev3.3s,fit_sev3.4s,fit_sev3.5s,fit_sev3.6s)
```

## Transect level severity without CPTA n = 183 ##### 
```{r transect_sev_all}
fit_sev1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|Region)+(1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
summary(fit_sev1)
#Region is also close to zero
fit_sev2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      Region+(1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
drop1(fit_sev2)
#drop region
fit_sev4 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
summary(fit_sev3)
# drop1(fit_sev3)
# that's it for model selection
# check residuals
# Es.sim <- simulateResiduals(fit_sev3)
# plot(Es.sim)
# plot(Es.sim$scaledResiduals~sevT$sBladeArea)
# plot(Es.sim$scaledResiduals~sevT$sDensityShoots)
# plot(Es.sim$scaledResiduals~as.factor(sevT$TidalHeight))
# plot(Es.sim$scaledResiduals~sevT$Region)
performance(fit_sev3)
icc(fit_sev3, by_group = TRUE)

 ### different way to do this is to look at AIC of specific models
 # remove blade area : tidal height interaction
 fit_sev4.1 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 
 # remove shoot density : tidal height interaction
fit_sev4.2 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
  # remove epiphyte load : tidal height interaction
fit_sev4.3 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 # no interactions
 fit_sev4.4 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 # no blade area
 fit_sev4.5 <- glmmTMB(Severity~sDensityShoots+sEpiphytePerAreaMean+TidalHeight+
                      sDensityShoots:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 # no shoot density
 fit_sev4.6 <- glmmTMB(Severity~sBladeArea+sEpiphytePerAreaMean+TidalHeight+
                      sBladeArea:TidalHeight+sEpiphytePerAreaMean:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 # no epiphyte load
 fit_sev4.7 <- glmmTMB(Severity~sBladeArea+sDensityShoots+TidalHeight+
                      sBladeArea:TidalHeight+sDensityShoots:TidalHeight+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 # no tidal height (or interactions)
 fit_sev4.8 <- glmmTMB(Severity~sBladeArea+sDensityShoots+sEpiphytePerAreaMean+
                      (1|MeadowId),
                    data=sevT,
                    family=beta_family(link = "logit"))
 AIC(fit_sev4,fit_sev4.1,fit_sev4.2,fit_sev4.3,fit_sev4.4,
     fit_sev4.5,fit_sev4.6,fit_sev4.7,fit_sev4.8)
```